---
title: "01_dataset_creation"
output: html_document
---

# Environment

```{r}
library(data.table)
library(summarytools)
library(zoo)
library(dplyr)
library(magrittr)
library(lubridate)
library(dtplyr)
library(vroom)
library(stringi)
library(tidyr)
library(parallel)

# defining functions

special_date = function(number){
  date = if_else(number%in%c(-1,0,NA), as.Date(NA), as.Date(as.character(number),format="%Y%m%d"))
  return(date)
}

convertir_valores <- function(x) {ifelse(x == -1, 0, 1)}
```


# Data Load

## poblacion_diana
Se filtra el sexo en dos posibles categorías: hombre y mujer. Otros eran menor al 0.1%.

```{r}
poblacion_diana <- fread("/opt/datos_compartidos/vacunas_data/total/ENC_VACUNAS_POB_DIANA.txt", 
                         header = T, 
                         encoding = 'Latin-1')

poblacion_diana$COD_FEC_NACIMIENTO <- as.Date(as.character(poblacion_diana$COD_FEC_NACIMIENTO), format = "%Y%m%d") 

# filtering in two sexes.
poblacion_diana <- poblacion_diana %>% as_tibble()
poblacion_diana <- poblacion_diana %>% 
  dplyr::filter(DESC_SEXO == "Hombre" |
           DESC_SEXO == "Mujer")

# changing reference in sex
poblacion_diana$DESC_SEXO <- relevel(as.factor(poblacion_diana$DESC_SEXO), ref = "Mujer" )

# age at 2020-03-01 field
poblacion_diana$age<-trunc((poblacion_diana$COD_FEC_NACIMIENTO %--% '2020-03-01' ) / years(1))

# mortality field
poblacion_diana$mortality <- convertir_valores(poblacion_diana$FEC_FALLECIMIENTO)
```
## pupa

Periodo-Usuario PAtologías

```{r}
pupa <- fread("/opt/datos_compartidos/vacunas_data/total/ENC_VACUNAS_PUPA.txt",header = T,encoding = 'Latin-1');pupa<-as.data.frame(pupa);

# Remove duplicates
pupa %<>% distinct()

# dates need to be converted from the original source, missing means NO
pupa$COD_FEC_INI_PATOLOGIA <- as.Date(as.character(pupa$COD_FEC_INI_PATOLOGIA), format="%Y%m%d")

# First, we convert the categorical variables to one without capital letters and accents
pupa$DESC_PATOLOGIA_CRONICA <- tolower(make.names(stri_trans_general(pupa$DESC_PATOLOGIA_CRONICA, "latin-ascii")))



#TODO, FILTER by admission or infection date?
# Join the two data sets by the id column
# df_joined <- df1 %>%
#   left_join(df2, by = "id")
# 
# # Filter the joined data frame to select rows where the date difference between the two date columns is 59 days
# df_filtered <- df_joined %>%
#   filter(abs(as.numeric(difftime(date1, date2, units = "days"))) == 60)
# 
# # Select only the categorical column from the filtered data frame
# df_final <- df_filtered %>%
#   select(categorical_column)

#This will create a new column for each unique element in the DESC_PATOLOGIA_CRONICA column, and fill the cells with a value of 1 for rows where the DESC_PATOLOGIA_CRONICA value matches the column name, and 0 for all other rows.

# Convert the data frame to a data.table since dplyr is not efficent enough
pupa_dt <- setDT(pupa)

# Group by and mutate in a single step
pupa_opt <- pupa_dt[, value := 1, by = c("NUHSA_ENCRIPTADO", "DESC_PATOLOGIA_CRONICA")]

# Spread in a single step
pupa_opt <- dcast(pupa_opt, NUHSA_ENCRIPTADO ~ DESC_PATOLOGIA_CRONICA, value.var = "value", fill = 0)



# Replace the values in the original data frame with 1 if any value in the row is greater than 0, and 0 otherwise

# Select the first column
df_first <- pupa_opt[, 1]

# Select the columns starting from the second column
df_selected <- pupa_opt[, 2:ncol(pupa_opt)]

# Replace the values in the selected columns with 1 if they are greater than 0, and 0 otherwise
df_selected <- apply(df_selected, 2, function(x) ifelse(x > 0, 1, 0))

# Append the transformed columns to the first column
pupa_opt <- cbind(df_first, df_selected)

remove(df_first); remove(df_selected)
```

## periodos pandemicos

Información extraída del "Informe COVID-19 en Andalucía durante la fase aguda de la pandemia", elaborado por el Servicio de Estadísticas Sanitarias. Consejería de Salud y Familias. Disponible en el siguiente [enlace (exige red corporativa)](https://www.ieca.junta-andalucia.es/institutodeestadisticaycartografia/badea/operaciones/consulta/anual/50147?CodOper=b3_2314&codConsulta=50147).

```{r}
# ondas_pandemicas_andalucia <-
#   data.frame(
#     nombre_periodo = c("Periodo 1", "Periodo 2", "Periodo 3", "Periodo 4", "Periodo 5", "Periodo 6", "Periodo 7"),
#     fecha_inicio = c("2020-01-01", "2020-05-11", "2020-12-21", "2021-03-08", "2021-06-21", "2021-10-11", "2022-03-28"),
#     fecha_fin = c("2020-05-10", "2020-12-20", "2021-03-07", "2021-06-20", "2021-10-10", "2022-03-27", "2022-11-01"))
# 
# 
# write.csv2(
#   ondas_pandemicas_andalucia,
#   file = "/opt/datos_compartidos/vacunas_data/05_Periodos_pandemicos.csv",
#   row.names = FALSE)


periodos_pandemicos_andalucia <- 
  read.csv2("/opt/datos_compartidos/vacunas_data/05_Periodos_pandemicos.csv") %>% 
  mutate(nombre_periodo = factor(nombre_periodo,
                                 ordered = TRUE),
         fecha_inicio = as.Date(fecha_inicio),
         fecha_fin = as.Date(fecha_fin))
```

## resultados_pruebas_COVID

```{r RESULTADOS_PRUEBAS_COVID}
# '_t' means total population, second data load
# database importation and tidying
resultados_pruebas_COVID <-
  fread("/opt/datos_compartidos/vacunas_data/total/ENC_VACUNAS_TEST_COVID.txt",
        header = T,
        encoding = 'UTF-8') %>% 
  as.data.frame() %>% 
  select(-TIPO_PRUEBA) %>% 
  mutate(FECHA_MUESTRA = as.Date(as.character(FECHA_MUESTRA),
                                 format = "%d/%m/%Y"))
# infection count per person
infect_COVID_num <-
  resultados_pruebas_COVID %>%
  group_by(NUHSA_ENCRIPTADO) %>%
  summarise(infecc_covid_num = n())

# first COVID infection with date available
# TODO: pendientes de PRDA
infect_COVID_first <-
  resultados_pruebas_COVID %>%
  group_by(NUHSA_ENCRIPTADO) %>%
  filter(row_number() == 1 & DESC_T_PRUEBA == 'PCR' & !is.na(FECHA_MUESTRA)) %>% 
  rename(infect_COVID_first_date = FECHA_MUESTRA) %>%
  select(NUHSA_ENCRIPTADO, infect_COVID_first_date)
```

## episodios_cmbd

Episodios CMBD del 03/03/2019 al 01/04/2022.

One of the first 5 admission diagnoses, should be in this list

var1	freq	descripcion
U07.1	41192	COVID-19
J18.9	11869	Neumonía, microorganismo no especificado
J12.89	3384	Otros tipos de neumonía vírica
J22	3046	Infección aguda del tracto respiratorio inferior, no especificada
J96.00	2305	Insuficiencia respiratoria aguda no especificada si con hipoxia o con hipercapnia
J96.01	2131	Insuficiencia respiratoria aguda con hipoxia
J96.02	1317	Insuficiencia respiratoria aguda con hipercapnia
J96.20	1177	Insuficiencia respiratoria aguda y crónica no especificada si con hipoxia o con hipercapnia
J90	1113	Derrame pleural, no clasificable bajo otro concepto
J47.0	1045	Bronquiectasias con infección aguda de vías respiratorias bajas
J96.21	772	Insuficiencia respiratoria aguda y crónica con hipoxia
J98.01	597	Broncoespasmo agudo
J18.0	512	Bronconeumonía, microorganismo no especificado
J84.9	497	Enfermedad intersticial pulmonar, no especificada
J96.91	365	Insuficiencia respiratoria no especificada, con hipoxia
J96.90	348	Insuficiencia respiratoria no especificada, sin especificar si con hipoxia o con hipercapnia
J96.92	339	Insuficiencia respiratoria no especificada, con hipercapnia
J84.89	330	Otros tipos de enfermedades pulmonares intersticiales especificadas
J81.0	302	Edema agudo de pulmón
I26.99	291	Otra embolia pulmonar sin cor pulmonale agudo
J06.9	204	Infección respiratoria aguda del tracto respiratorio superior, no especificada
J98.9	112	Trastorno respiratorio, no especificado
J95.89	107	Otras complicaciones y trastornos posprocedimiento de aparato respiratorio, no clasificados bajo otro concepto
J12.9	86	Neumonía vírica, no especificada
O98.52	82	Otras enfermedades virales que complican el parto
J84.111	72	Neumonía intersticial idiopática, no especificada de otro modo
O98.513	70	Otras enfermedades virales que complican el embarazo, tercer trimestre
J20.8	66	Bronquitis aguda por otros organismos especificados
J96.11	64	Insuficiencia respiratoria crónica con hipoxia
J00	62	Nasofaringitis aguda [resfriado común]
J12.81	42	Neumonía por coronavirus asociado al SARS
J84.114	42	Neumonitis intersticial aguda
J39.8	41	Otras enfermedades especificadas del tracto respiratorio superior
J84.115	15	Enfermedad pulmonar intersticial asociada a bronquiolitis respiratoria

TODO hay 92k pacientes sin fecha de ingreso. rafa: a mi ma salen todos con Fecha ingreso y alta.
TODO puede ser que ya estuvieran ingresados.

```{r}
icd_ofinterest <- c('U07.1', 'J18.9', 'J12.89', 'J22', 'J96.00', 'J96.01', 'J96.02', 'J96.20', 'J90', 'J47.0', 'J96.21', 'J98.01', 'J18.0', 'J84.9', 'J96.91', 'J96.90', 'J96.92', 'J84.89', 'J81.0', 'I26.99', 'J06.9', 'J98.9', 'J95.89', 'J12.9', 'O98.52', 'J84.111', 'O98.513', 'J20.8', 'J96.11', 'J00', 'J12.81', 'J84.114', 'J39.8', 'J84.115', 'J84.113', 'B34.2')
episodios_cmbd <- fread("/opt/datos_compartidos/vacunas_data/total/ENC_VACUNAS_CMBD.txt", header = T, encoding = "Latin-1") %>% 
  as.data.frame()
cmbd_t <- as.data.frame(episodios_cmbd)

# TODO no hay fecha de ingreso en UCI. No podemos calcular dias UCIA. Mirar traslados. rafa: no creo que tengamos acceso a esto.

# dates need to be converted from the original source, missing means NO
episodios_cmbd$COD_FEC_INGRESO <- as.Date(as.character(episodios_cmbd$COD_FEC_INGRESO),format="%Y%m%d")
episodios_cmbd$COD_FEC_ALTA <- as.Date(as.character(episodios_cmbd$COD_FEC_ALTA),format="%Y%m%d")

# dias estancia hospitalaria
episodios_cmbd <- episodios_cmbd %>% 
  mutate(estancia_hospitalaria_dias = as.numeric(COD_FEC_ALTA - COD_FEC_INGRESO))

# dias entre infeccion e ingreso

# joining  the df containing infect_COVID_first_date

result <- left_join(episodios_cmbd, infect_COVID_first)

result %<>% 
  mutate(infeccion_ingreso_dias = infect_COVID_first_date - COD_FEC_INGRESO)

#  select only cols from names(episodios_cmbd)
episodios_cmbd<-result %>% select(NUHSA_ENCRIPTADO,COD_FEC_INGRESO,COD_FEC_ALTA,COD_TIPO_INGRESO,DESC_TIPO_INGRESO,COD_TIPO_ALTA,DESC_TIPO_ALTA,COD_D1,COD_D2,COD_D3,COD_D4,COD_D5,COD_D6,COD_D7,COD_D8,COD_D9,COD_D10,COD_D11,COD_D12,COD_D13,COD_D14,COD_D15,COD_D16,COD_D17,COD_D18,COD_D19,COD_D20,COD_P1,COD_P2,COD_P3,COD_P4,COD_P5,COD_P6,COD_P7,COD_P8,COD_P9,COD_P10,COD_P11,COD_P12,COD_P13,COD_P14,COD_P15,COD_P16,COD_P17,COD_P18,COD_P19,COD_P20,NUM_DIAS_UCI,estancia_hospitalaria_dias,infeccion_ingreso_dias)

# estancia uci si/no
episodios_cmbd %<>% 
mutate(icu_stay_bin = as.factor(if_else((NUM_DIAS_UCI == 0 | is.na(NUM_DIAS_UCI)),0,1)))

# dias entre infeccion e ingreso hospitalario ha de ser de 30 días máximo. si no, no lo considero ingreso hospitalario por COVID
#los aún ingresados en el período de análisis habrían de tenerse en cuenta. Es menor de 30 si sale negativo también se cuenta.
episodios_cmbd <- episodios_cmbd %>% 
  mutate(hosp_stay_bin = if_else(infeccion_ingreso_dias <= 30, 1 ,0))

#we are just going to take into account first hospitalization
#filtramos los 5 códigos principales sólo por ICD relacionados con COVID, según conteo
# we are only selecting fist admission per patient
episodios_cmbd_first <- episodios_cmbd %>%
  select(
    NUHSA_ENCRIPTADO,
    COD_FEC_INGRESO,
    COD_FEC_ALTA,
    COD_TIPO_ALTA,
    DESC_TIPO_ALTA,
    hosp_stay_bin,
    estancia_hospitalaria_dias,
    icu_stay_bin,
    NUM_DIAS_UCI,
    COD_D1,
    COD_D2,
    COD_D3,
    COD_D4,
    COD_D5,
    infeccion_ingreso_dias
  ) %>%
  filter(hosp_stay_bin == 1) %>%
  filter(
    COD_D1 %in% icd_ofinterest | COD_D2 %in% icd_ofinterest | COD_D3 %in% icd_ofinterest | COD_D4 %in% icd_ofinterest | COD_D5 %in% icd_ofinterest
  ) %>%
  group_by(NUHSA_ENCRIPTADO) %>%
  filter(row_number() == 1) %>%
  select(
    NUHSA_ENCRIPTADO,
    COD_FEC_INGRESO,
    COD_FEC_ALTA,
    estancia_hospitalaria_dias,
    hosp_stay_bin,
    icu_stay_bin,
    infeccion_ingreso_dias
  )
```

## vacunas

Jaime M

TODO address non european shots: grepl('COVID-19 Sinopharm|COVID-19 Sputnik V|COVID-19 Covishield|COVID-19 CoronaVac (Sinovac Biotech)|COVID-19 CanSinoBIO|COVID-19 Covaxin|COVID-19 HIPRA')

Criterios de exclusión: pacientes de 5 dosis.

-	Vacunas de la gripe de las campañas 2020-21 y 2021-2022.
-	Vacunas de neumococo en cualquier momento hasta el 31/03/2022.
-	Todas las vacunas relacionadas con el COVID 19 hasta el 31/03/2022.

TODO STAGI: *Pedir campañas 2019-20.* 

TODO hay varias vacunas de neumo y gripe por paciente.
TODO estudiar numero de infecciones
```{r}
vacunas_t <- vroom(
  "/opt/datos_compartidos/vacunas_data/total/ENC_VACUNAS_VACUNAS_TODO.txt"
) %>%
  lazy_dt() %>%
  as.data.frame() %>% 
  left_join(.,infect_COVID_first,by = "NUHSA_ENCRIPTADO") %>% 
  mutate(
    FEC_VACUNACION = ymd(FEC_VACUNACION), # dates need to be converted from the original source, missing means NO
    vac_infecc_umbral = trunc((FEC_VACUNACION %--% infect_COVID_first_date) / days(1)) # dias entre vacunacion e infeccion
  )
# what shot was first per patient?
vacunas_t_covid_first <- vacunas_t %>%
  filter(vac_infecc_umbral >= 14) %>% # los dias entre la vacunacion e infeccion han de ser >= 14
  filter(
    grepl(
      'COVID-19 Pfizer/BioNTech|COVID-19 Moderna|COVID-19 AstraZeneca|COVID-19 Janssen|COVID-19 Novavax',
      NOM_VACUNA
    )
  ) %>%
  group_by(NUHSA_ENCRIPTADO) %>%
  arrange(FEC_VACUNACION) %>% 
  filter(row_number() == 1) %>% 
  transmute(
    NUHSA_ENCRIPTADO,
    covid_first_name = NOM_VACUNA,
    covid_first_date = FEC_VACUNACION
  )
# TODO cómo afecta el umbral de 14 días a la pauta completa?
# full shots
vacunas_t_covid_full <- vacunas_t %>%
  filter(
    grepl(
      'COVID-19 Pfizer/BioNTech|COVID-19 Moderna|COVID-19 AstraZeneca|COVID-19 Janssen|COVID-19 Novavax',
      NOM_VACUNA
    )
  ) %>%
  group_by(NUHSA_ENCRIPTADO) %>%
  arrange(FEC_VACUNACION) %>% 
  filter(row_number()==3) %>% 
  transmute(
    NUHSA_ENCRIPTADO,
    covid_full_name = NOM_VACUNA,
    covid_full_date = FEC_VACUNACION
  )

# number of shots ?
vacunas_t_covid_pautas <- vacunas_t %>%
  group_by(NUHSA_ENCRIPTADO) %>%
  filter(
    grepl(
      'COVID-19 Pfizer/BioNTech|COVID-19 Moderna|COVID-19 AstraZeneca|COVID-19 Janssen|COVID-19 Novavax',
      NOM_VACUNA
    )
  ) %>% 
  summarise(pautas_covid = n())



# neumococo simple version, was the patient vaccinated? (bin)
vacunas_neumococo <- vacunas_t %>%
  filter(vac_infecc_umbral >= 14)%>%
  filter(
    grepl('Neumococo', NOM_VACUNA)
  ) %>% 
  transmute(
    NUHSA_ENCRIPTADO,
    vacuna_neumo = 1
  ) %>% 
  distinct()
# gripe simple version, was the patient vaccinated? (bin)
vacunas_gripe <- vacunas_t %>%
  filter(vac_infecc_umbral >= 14)%>%
  filter(
    grepl('Gripe', NOM_VACUNA)
  ) %>% 
  transmute(
    NUHSA_ENCRIPTADO,
    vacuna_gripe = 1
  ) %>% 
  distinct()

vacunas_inmunity <- vacunas_t %>%
  filter(vac_infecc_umbral >= 14)%>%
  filter(
    grepl('Gripe', NOM_VACUNA)
  ) %>% 
  transmute(
    NUHSA_ENCRIPTADO,
    vacuna_gripe = 1
  ) %>% 
  distinct()
# TODO hay varias vacunas de neumo y gripe por paciente.
```

# Final dataset

## Joining across datasets

```{r}
print(Sys.time())
estudio_vacunas_t_final<-Reduce(function(...) merge(..., all.x=TRUE), list(
    infect_COVID_first
  , poblacion_diana #select(NUHSA_ENCRIPTADO,DESC_SEXO,age, mortality)
  , pupa_opt
  , infect_COVID_num  
  , episodios_cmbd_first # select(NUHSA_ENCRIPTADO,estancia_hospitalaria_dias,hosp_stay_bin,icu_stay_bin,infeccion_ingreso_dias) 
  , vacunas_t_covid_first
  , vacunas_t_covid_full
  , vacunas_t_covid_pautas
))
print(Sys.time())

saveRDS(estudio_vacunas_t_final, file = "estudio_vacunas_t_final.rds")
```

## Selecting only required columns

```{r}
cols_to_exclude <- c("NUHSA_ENCRIPTADO", "COD_FEC_NACIMIENTO")
cols_to_include <- setdiff(names(estudio_vacunas_t_final), cols_to_exclude)
estudio_vacunas_t_final_selected<-estudio_vacunas_t_final %>% select(cols_to_include)
```


## Adressing missingness

### Mapping initial missingness

```{r}
# Count the number of missing values in each column of the data frame `df`
miss_count <- colSums(is.na(estudio_vacunas_t_final_selected))

# Calculate the percentage of missing values in each column
miss_percent <- round(miss_count / nrow(estudio_vacunas_t_final_selected) * 100,2)

# Order the results by the column names in the original data frame
ordered_miss_percent <- miss_percent[order(names(estudio_vacunas_t_final_selected))]

# Save the results in a new data frame
miss_percent_df <- data.frame(Column = names(ordered_miss_percent), Missing_Percent = ordered_miss_percent)
```

### Columns NA filling

```{r}
# except for date fields, for every NA cell, we replace it with 0 since
# for cat variables empty means NO and for days related cols, NA means 0 days.

fill_cols_to_exclude <- c("COD_FEC_ALTA", "COD_FEC_INGRESO","covid_full_date","covid_first_date")
fill_cols_to_include <- setdiff(names(estudio_vacunas_t_final_selected), fill_cols_to_exclude)

# Replace NAs with 0s
estudio_vacunas_t_final_selected[, fill_cols_to_include][is.na(estudio_vacunas_t_final_selected[, fill_cols_to_include])] <-'0'
```

### Mapping final missingness

```{r}
# Count the number of missing values in each column of the data frame `df`
miss_count <- colSums(is.na(estudio_vacunas_t_final_selected))

# Calculate the percentage of missing values in each column
miss_percent <- round(miss_count / nrow(estudio_vacunas_t_final_selected) * 100,2)

# Order the results by the column names in the original data frame
ordered_miss_percent <- miss_percent[order(names(estudio_vacunas_t_final_selected))]

# Save the results in a new data frame
miss_percent_df <- data.frame(Column = names(ordered_miss_percent), Missing_Percent = ordered_miss_percent)
```


## New variables creation

```{r}
estudio_vacunas_t_final_selected<-estudio_vacunas_t_final_selected%>%
  
 mutate(cancer_organo_solido = +(
                                  cancer.de.bronquio.y.pulmon | 
                                  cancer.de.cabeza.y.cuello |
                                  cancer.de.estomago |
                                  cancer.de.hueso.y.tejidos.blandos |
                                  cancer.de.mama |
                                  cancer.de.ovario |
                                  cancer.de.pancreas |
                                  cancer.de.prostata |
                                  cancer.de.rinon.y.pelvis.renal |
                                  cancer.de.testiculo |
                                  cancer.de.tiroides |
                                  cancer.de.utero |
                                  cancer.de.vejiga |
                                  cancer.de.cuello.uterino |
                                  melanoma.de.piel),
        
         cancer_hematologico = +( 
                                  cancer.inmunoproliferativo |
                                  enfermedad.de.hodgkin |
                                  leucemia |
                                  linfoma.no.hodgkin |
                                  sarcoma.de.kaposi)
        )

vacunas_t_covid_pautas %<>% 
  mutate(pautas_covid = as.factor(pautas_covid))

```

(Coalesce de fallecimiento y fecha de ingreso)

# Exclusion criteria

No exclusion criteria by now.

```{r}
print('Número pacientes con infección COVID confirmada por PCR y fecha de infección disponible')
a<-nrow(estudio_vacunas_t_final_selected)
a

# estudio_vacunas_final_selected<-estudio_vacunas_final%>%filter(pautas_covid <5 )
# print('Pacientes excluídos con más de 4 dosis')
# b<-nrow(estudio_vacunas_t_final_selected)
# a-b

print('Número final de pacientes')
nrow(estudio_vacunas_t_final_selected)

saveRDS(estudio_vacunas_t_final_selected, file = "estudio_vacunas_t_final_selected.rds")

# feather format is compatible both with R and python
write_feather(estudio_vacunas_t_final_selected, "estudio_vacunas_t_final_selected.feather")
```

