---
title: "01_dataset_creation"
output: html_document
---

# Environment

```{r}
library(survival)
library(survminer)
library(forestmodel)
library(skimr)
library(mgcv)
library(randomForest)
library(parallel)
library(parglm)
library(pROC)
library(ggplot2)
library(feather)

# metric performance

precision <- function(matrix) {
	# True positive
    tp <- matrix[2, 2]
	# false positive
    fp <- matrix[1, 2]
    return (tp / (tp + fp))
}

recall <- function(matrix) {
# true positive
    tp <- matrix[2, 2]# false positive
    fn <- matrix[2, 1]
    return (tp / (tp + fn))
}


```

# survival model

```{r}
#TODO evaluate efficiency
# event as numeric
estudio_vacunas_final_selected$estancia_hospitalaria<-as.numeric(estudio_vacunas_final_selected$estancia_hospitalaria)

res.cox <- coxph(Surv(estancia_hospitalaria_dias, estancia_hospitalaria) ~ DESC_SEXO + RESIDE_RESIDENCIA + GRUPO_7 + age + cancer + VIH + EPOC + ASMA + INSUF_CARDIACA + CARDIOPATÍA_ISQUÉMCIA + HEPATOPATIA_CRÓNICA + CIRROSIS + INSUF_RENAL_CRÓNICA + vac_groupB, data =  estudio_vacunas_final_selected)
```


## Plotting

```{r}
ggsurvplot(res.cox, data = estudio_vacunas_final_selected)
```

# logreg hosp_stay_bin

## Fitting the model

```{r}
print('glm')
print(Sys.time())
logreg_hosp_stay_bin <- glm(hosp_stay_bin ~ DESC_SEXO+ age+ asma +cancer.colorrectal+ cancer.de.bronquio.y.pulmon+cancer.de.cabeza.y.cuello + cancer.de.cuello.uterino +cancer.de.estomago+cancer.de.higado.y.vias.biliares+cancer.de.hueso.y.tejidos.blandos+cancer.de.mama+cancer.de.ovario+cancer.de.pancreas+cancer.de.prostata+cancer.de.rinon.y.pelvis.renal+cancer.de.testiculo+cancer.de.tiroides+cancer.de.utero+cancer.de.vejiga+cancer.inmunoproliferativo+cardiopatia.isquemica+cirrosis.hepatica+diabetes+enfermedad.de.hodgkin+epoc+hepatopatia.cronica.excepto.cirrosis+hipertension+insuficiencia.cardiaca+insuficiencia.renal.cronica+leucemia+linfoma.no.hodgkin+melanoma.de.piel+sarcoma.de.kaposi+vih+infecc_covid_num, pautas_covid,
data =  estudio_vacunas_t_final_selected,family=binomial() )
print(Sys.time())
```
## Summary of the model

```{r}
summary(logreg_hosp_stay_bin)
```

## Odds ratio

```{r fig.height=16, fig.width=10}
print(Sys.time())
print(forest_model(logreg_hosp_stay_bin))
print(Sys.time())
```
## Odds ratio II

```{r}
library(parallel)

# Create a cluster with two cores for parallel processing
cl <- makeCluster(2)

# Wrap the calculation of predictions and confidence intervals in a function
calculate <- function(logreg_hosp_stay_bin, estudio_vacunas_t_final_selected) {
  pred_probs <- predict(logreg_hosp_stay_bin, estudio_vacunas_t_final_selected, type="response")
  conf_int <- confint(logreg_hosp_stay_bin, level=0.95)
  
  list(pred_probs, conf_int)
}

# Split the data into two chunks and send one chunk to each core for processing
split_data <- split(estudio_vacunas_t_final_selected, rep(1:2, length.out = nrow(estudio_vacunas_t_final_selected)))
results <- parSapply(cl, split_data, calculate, logreg_hosp_stay_bin = logreg_hosp_stay_bin)

# Combine the results from the two cores
pred_probs_combined <- do.call(rbind, lapply(results, function(x) x[[1]]))
conf_int_combined <- do.call(rbind, lapply(results, function(x) x[[2]]))

# Convert the results to a data frame
results_df <- data.frame(variable = names(logreg_hosp_stay_bin$coefficients),
                         coef = logreg_hosp_stay_bin$coefficients,
                         lower = conf_int_combined[,1],
                         upper = conf_int_combined[,2])

# Plot the forest plot
ggplot(results_df, aes(x=reorder(variable, coef), ymin=lower, ymax=upper, y=coef)) +
  geom_pointrange(alpha=0.7) + 
  geom_hline(yintercept=0, color="red") +
  xlab("Variable") + 
  ylab("Coeficiente (IC 95%)") + 
  ggtitle("Forest plot del modelo GLM") + 
  coord_flip()

# Stop the cluster
stopCluster(cl)
```


## Model Performance

```{r fig.height=5, fig.width=2}
y_test<-estudio_vacunas_final_selected$hosp_stay_bin
  
predict <- predict(logreg_hosp_stay_bin, estudio_vacunas_final_selected, type = 'response')
# confusion matrix
table_mat <- table(y_test, predict > 0.85)
table_mat

accuracy <- sum(diag(table_mat)) / sum(table_mat)
accuracy

prec <- precision(table_mat)
prec

rec <- recall(table_mat)
rec

f1 <- 2 * ((prec * rec) / (prec + rec))
f1

#AUTOC plot

predicted_bin<-as.numeric(c(predict > 0.1))
rocobj <- roc(y_test, predicted_bin)

auc <- round(auc(y_test, predicted_bin),2)

#create ROC plot
ggroc(rocobj, colour = 'steelblue', size = 2) +
  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')'))

metrics<-round(as.data.frame(rbind(accuracy,auc)),2)
metrics$metric<-rownames(metrics)
metrics$metric<-toupper(metrics$metric)

names(metrics)<-c('Value', 'Metric')

ggplot(data=metrics, aes(x=Metric, y=Value)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Value), vjust=1.6, color="white", size=3.5)+
  theme_minimal()
```

# random forest

## Fitting the model

```{r}
# Selecciona número de núcleos para paralelización
ncores <- detectCores() - 1
cl <- makeCluster(ncores)
clusterExport(cl, c("estudio_vacunas_t_final_selected", "hosp_stay_bin"))

# Crear modelo de Random Forest
rf_model <- foreach(ntree = rep(500, ncores), .combine = combine, .packages = "randomForest") %dopar% {
  set.seed(123)
  randomForest(hosp_stay_bin ~ DESC_SEXO+ age+ asma +cancer.colorrectal+ cancer.de.bronquio.y.pulmon+cancer.de.cabeza.y.cuello + cancer.de.cuello.uterino +cancer.de.estomago+cancer.de.higado.y.vias.biliares+cancer.de.hueso.y.tejidos.blandos+cancer.de.mama+cancer.de.ovario+cancer.de.pancreas+cancer.de.prostata+cancer.de.rinon.y.pelvis.renal+cancer.de.testiculo+cancer.de.tiroides+cancer.de.utero+cancer.de.vejiga+cancer.inmunoproliferativo+cardiopatia.isquemica+cirrosis.hepatica+diabetes+enfermedad.de.hodgkin+epoc+hepatopatia.cronica.excepto.cirrosis+hipertension+insuficiencia.cardiaca+insuficiencia.renal.cronica+leucemia+linfoma.no.hodgkin+melanoma.de.piel+sarcoma.de.kaposi+vih+infecc_covid_num,
  data = estudio_vacunas_t_final_selected,
  ntree = ntree,
  importance = TRUE)
}

# Cerrar clúster
stopCluster(cl)

# Tomar el promedio de los modelos para tener un solo modelo final
rf_model_final <- mean(rf_model)

```

## Explaining the model

```{r}
library(ggplot2)
library(caret)

# calcular la importancia de las variables
imp <- varImp(rf_model, scale=FALSE)

# graficar la importancia de las variables
ggplot(imp, aes(x=reorder(variable, Overall), y=Overall)) +
  geom_bar(stat="identity") + 
  xlab("Variable") + 
  ylab("Importance") + 
  ggtitle("Importancia de las variables en el modelo Random Forest")

# predecir las probabilidades con el modelo Random Forest
pred_probs <- predict(rf_model, estudio_vacunas_t_final_selected, type="prob")

# graficar las decisiones del modelo
ggplot(estudio_vacunas_t_final_selected, aes(x=age, y=pred_probs[,2], color=hosp_stay_bin)) +
  geom_point() + 
  xlab("Edad") + 
  ylab("Probabilidad de permanencia en el hospital") + 
  ggtitle("Decisiones del modelo Random Forest")

```


# logreg_icu_stay_bin

## Fitting the model

```{r}
print('glm')
print(Sys.time())
logreg_icu_stay_bin <- glm(icu_stay_bin ~ age + DESC_SEXO + RESIDE_RESIDENCIA + GRUPO_7 + cancer + VIH + EPOC + ASMA + INSUF_CARDIACA + CARDIOPATÍA_ISQUÉMCIA + HEPATOPATIA_CRÓNICA + CIRROSIS + INSUF_RENAL_CRÓNICA + vac_COV_Neumo_Gripe, data =  estudio_vacunas_final_selected,family=binomial() )
print(Sys.time())
```

```{r}
summary(logreg_icu_stay_bin)
```

## Odds ratio

```{r fig.height=16, fig.width=10}
print(Sys.time())
print(forest_model(logreg_icu_stay_bin))
print(Sys.time())
```

## Model Performance

```{r fig.height=5, fig.width=2}
y_test<-estudio_vacunas_final_selected$icu_stay_bin
  
predict <- predict(logreg_icu_stay_bin, estudio_vacunas_final_selected, type = 'response')
# confusion matrix
table_mat <- table(y_test, predict > 0.1)
table_mat

accuracy <- sum(diag(table_mat)) / sum(table_mat)
accuracy

prec <- precision(table_mat)
prec

rec <- recall(table_mat)
rec

f1 <- 2 * ((prec * rec) / (prec + rec))
f1

#AUTOC plot

predicted_bin<-as.numeric(c(predict > 0.3))
rocobj <- roc(y_test, predicted_bin)

auc <- round(auc(y_test, predicted_bin),2)

#create ROC plot
ggroc(rocobj, colour = 'steelblue', size = 2) +
  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')'))

metrics<-round(as.data.frame(rbind(accuracy,auc)),2)
metrics$metric<-rownames(metrics)
metrics$metric<-toupper(metrics$metric)

names(metrics)<-c('Value', 'Metric')

ggplot(data=metrics, aes(x=Metric, y=Value)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Value), vjust=1.6, color="white", size=3.5)+
  theme_minimal()
```

#logreg_mortality

##Fitting the model
```{r}
print('glm')
print(Sys.time())
logreg_mortality <- glm(mortality ~ age + DESC_SEXO + RESIDE_RESIDENCIA + GRUPO_7 + cancer + VIH + EPOC + ASMA + INSUF_CARDIACA + CARDIOPATÍA_ISQUÉMCIA + HEPATOPATIA_CRÓNICA + CIRROSIS + INSUF_RENAL_CRÓNICA + vac_COV_Neumo_Gripe, data =  estudio_vacunas_final_selected,family=binomial() )
print(Sys.time())

```

##Summary the model

```{r}

summary(logreg_mortality)

```

##Odds ratio

```{r}

print(Sys.time())
print(forest_model(logreg_mortality))
print(Sys.time())

```

##Model Performance

````{r}

y_test<-estudio_vacunas_final_selected$icu_stay_bin
  
predict <- predict(logreg_mortality, estudio_vacunas_final_selected, type = 'response')

# confusion matrix
table_mat <- table(y_test, predict > 0.1)
table_mat

accuracy <- sum(diag(table_mat)) / sum(table_mat)
accuracy

prec <- precision(table_mat)
prec

rec <- recall(table_mat)
rec

f1 <- 2 * ((prec * rec) / (prec + rec))
f1

#AUTOC plot

predicted_bin<-as.numeric(c(predict > 0.3))
rocobj <- roc(y_test, predicted_bin)

auc <- round(auc(y_test, predicted_bin),2)

#create ROC plot
ggroc(rocobj, colour = 'steelblue', size = 2) +
  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')'))

metrics<-round(as.data.frame(rbind(accuracy,auc)),2)
metrics$metric<-rownames(metrics)
metrics$metric<-toupper(metrics$metric)

names(metrics)<-c('Value', 'Metric')

ggplot(data=metrics, aes(x=Metric, y=Value)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Value), vjust=1.6, color="white", size=3.5)+
  theme_minimal()
