---
title: "02_modelling"
output: html_document
---

# Environment

```{r}
library(survival)
library(survminer)

library(skimr)
library(mgcv)
library(parallel)
library(sjPlot)
library(parglm)
library(pROC)
library(ggplot2)

library(forestmodel)
library(gtsummary)
library(tidyverse)
library(janitor)

estudio_vacunas_t_final_selected <- readRDS(file = "/opt/datos_compartidos/vacunas_data/Rfiles/toberemoved/estudio_vacunas_t_final_selected_v13feb.rds")
estudio_vacunas_t_final_selected <- estudio_vacunas_t_final_selected %>% clean_names()

# estudio_vacunas_t_final_selected %>% str()
estudio_vacunas_t_final_selected <- estudio_vacunas_t_final_selected %>% 
  mutate(hosp_stay_bin = as.numeric(hosp_stay_bin))

# metric performance

precision <- function(matrix) {
	# True positive
    tp <- matrix[2, 2]
	# false positive
    fp <- matrix[1, 2]
    return (tp / (tp + fp))
}

recall <- function(matrix) {
# true positive
    tp <- matrix[2, 2]# false positive
    fn <- matrix[2, 1]
    return (tp / (tp + fn))
}


```

# Database Summary 
```{r}
tbl_summary(estudio_vacunas_t_final_selected, 
            by = hosp_stay_bin) %>% 
  # add_p() %>% 
  add_overall() %>% 
  bold_labels() 
  
```


# survival model

```{r}
#TODO evaluate efficiency
# event as numeric
estudio_vacunas_final_selected <- estudio_vacunas_final_selected %>% 
  mutate(estancia_hospitalaria = as.numeric(estancia_hospitalaria))

# retiro + RESIDE_RESIDENCIA + GRUPO_7 porque no está en la base de datos 
res.cox <- coxph(Surv(estancia_hospitalaria_dias, estancia_hospitalaria) ~ desc_sexo + age + cancer + vih + epoc + asma + insuficiencia_cardiaca + cardiopatia_isquemica + hepatopatia_cronica_excepto_cirrosis + cirrosis_hepatica + insuficiencia_renal_cronica + vac_groupB, data =  estudio_vacunas_final_selected)
```

## Plotting

```{r}
ggsurvplot(res.cox, data = estudio_vacunas_final_selected)
```


# logreg hosp_stay_bin

## Fitting the model

```{r}
# retiro + RESIDE_RESIDENCIA + GRUPO_7 porque no está en la base de datos 

print('glm')
print(Sys.time())
logreg_hosp_stay_bin <- glm(
  hosp_stay_bin ~ age_num_decada + desc_sexo + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc + hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + pautas_covid,
  data =  estudio_vacunas_t_final_selected, 
  family=binomial())
print(Sys.time())
```
## Summary of the model

```{r}
summary(logreg_hosp_stay_bin)
```

## Odds ratio

```{r fig.height=16, fig.width=10}
print(Sys.time())
# tbl_regression(logreg_hosp_stay_bin, exponentiate = TRUE) %>%
#   bold_labels() %>%
#   italicize_levels()
print(forest_model(logreg_hosp_stay_bin))
print(Sys.time())
```

## Model Performance

```{r fig.height=5, fig.width=2}
y_test <- estudio_vacunas_final_selected$hosp_stay_bin
  
predict <- predict(logreg_hosp_stay_bin, estudio_vacunas_final_selected, type = 'response')
# confusion matrix
table_mat <- table(y_test, predict > 0.85)
table_mat

accuracy <- sum(diag(table_mat)) / sum(table_mat)
accuracy

prec <- precision(table_mat)
prec

rec <- recall(table_mat)
rec

f1 <- 2 * ((prec * rec) / (prec + rec))
f1

#AUTOC plot

predicted_bin <- as.numeric(c(predict > 0.1))
rocobj <- roc(y_test, predicted_bin)

auc <- round(auc(y_test, predicted_bin),2)

#create ROC plot
ggroc(rocobj, colour = 'steelblue', size = 2) +
  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')'))

metrics <- round(as.data.frame(rbind(accuracy,auc)),2)
metrics$metric <- rownames(metrics)
metrics$metric <- toupper(metrics$metric)

names(metrics) <- c('Value', 'Metric')

ggplot(data=metrics, aes(x=Metric, y=Value)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Value), vjust=1.6, color="white", size=3.5)+
  theme_minimal()
```


# logreg_icu_stay_bin

## Fitting the model

```{r}
# retiro + RESIDE_RESIDENCIA + GRUPO_7 + vac_COV_Neumo_Gripe porque no está en la base de datos 

print('glm')
print(Sys.time())
logreg_icu_stay_bin <- glm(
  icu_stay_bin ~ age + desc_sexo + cancer + vih + epoc + asma + insuficiencia_cardiaca + cardiopatia_isquemica + hepatopatia_cronica_excepto_cirrosis + cirrosis_hepatica + insuficiencia_renal_cronica , 
  data =  estudio_vacunas_final_selected,
  family=binomial() )
print(Sys.time())
```

```{r}
summary(logreg_icu_stay_bin)
```

## Odds ratio

```{r fig.height=16, fig.width=10}
print(Sys.time())
print(forest_model(logreg_icu_stay_bin))
print(Sys.time())
```

## Model Performance

```{r fig.height=5, fig.width=2}
y_test<-estudio_vacunas_final_selected$icu_stay_bin
  
predict <- predict(logreg_icu_stay_bin, estudio_vacunas_final_selected, type = 'response')
# confusion matrix
table_mat <- table(y_test, predict > 0.1)
table_mat

accuracy <- sum(diag(table_mat)) / sum(table_mat)
accuracy

prec <- precision(table_mat)
prec

rec <- recall(table_mat)
rec

f1 <- 2 * ((prec * rec) / (prec + rec))
f1

#AUTOC plot

predicted_bin<-as.numeric(c(predict > 0.3))
rocobj <- roc(y_test, predicted_bin)

auc <- round(auc(y_test, predicted_bin),2)

#create ROC plot
ggroc(rocobj, colour = 'steelblue', size = 2) +
  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')'))

metrics<-round(as.data.frame(rbind(accuracy,auc)),2)
metrics$metric<-rownames(metrics)
metrics$metric<-toupper(metrics$metric)

names(metrics)<-c('Value', 'Metric')

ggplot(data=metrics, aes(x=Metric, y=Value)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Value), vjust=1.6, color="white", size=3.5)+
  theme_minimal()
```

#logreg_mortality

##Fitting the model
```{r}
print('glm')
print(Sys.time())
logreg_mortality <- glm(
  mortality ~ age + desc_sexo + RESIDE_RESIDENCIA + GRUPO_7 + cancer + vih + epoc + asma + insuficiencia_cardiaca + cardiopatia_isquemica + hepatopatia_cronica_excepto_cirrosis + cirrosis_hepatica + insuficiencia_renal_cronica + vac_COV_Neumo_Gripe, 
  data =  estudio_vacunas_final_selected,
  family=binomial() )
print(Sys.time())

```

##Summary the model

```{r}

summary(logreg_mortality)

```

##Odds ratio

```{r}

print(Sys.time())
print(forest_model(logreg_mortality))
print(Sys.time())

```

##Model Performance

````{r}

y_test<-estudio_vacunas_final_selected$icu_stay_bin
  
predict <- predict(logreg_mortality, estudio_vacunas_final_selected, type = 'response')

# confusion matrix
table_mat <- table(y_test, predict > 0.1)
table_mat

accuracy <- sum(diag(table_mat)) / sum(table_mat)
accuracy

prec <- precision(table_mat)
prec

rec <- recall(table_mat)
rec

f1 <- 2 * ((prec * rec) / (prec + rec))
f1

#AUTOC plot

predicted_bin<-as.numeric(c(predict > 0.3))
rocobj <- roc(y_test, predicted_bin)

auc <- round(auc(y_test, predicted_bin),2)

#create ROC plot
ggroc(rocobj, colour = 'steelblue', size = 2) +
  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')'))

metrics<-round(as.data.frame(rbind(accuracy,auc)),2)
metrics$metric<-rownames(metrics)
metrics$metric<-toupper(metrics$metric)

names(metrics)<-c('Value', 'Metric')

ggplot(data=metrics, aes(x=Metric, y=Value)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Value), vjust=1.6, color="white", size=3.5)+
  theme_minimal()