---
title: "02b_modeling_improved"
# output: html_document
---

# Environment

```{r}
library(magrittr)
library(tableone)
library(kableExtra)
library(gtsummary)
library(rio)
library(summarytools)
library(zoo)
library(lubridate)
library(vroom)
library(stringr)
library(stringi)
library(tidyr)
library(janitor)
# library(skimr)
library(mgcv)
library(parallel)
library(parglm)
library(pROC)
library(ggplot2)
library(tibble)
library(dplyr)
library(feather)
library(broom)
library(htmlTable)
library(purrr)
library(parsnip)
library(recipes)
library(workflows)



estudio_vacunas_t_final_selected <- 
  readRDS("/opt/datos_compartidos/vacunas_data/Rfiles/estudio_vacunas_t_final_selected_manuscript.rds")



comorb_no_cancer <- estudio_vacunas_t_final_selected %>% 
  select(asma, cardiopatia_isquemica:diabetes, epoc:insuficiencia_renal_cronica, vih) %>% 
  colnames(.)

solid_cancers_vector <- estudio_vacunas_t_final_selected %>% 
  select(cancer_colorrectal:cancer_de_vejiga, melanoma_de_piel, sarcoma_de_kaposi) %>% 
  colnames(.)

hemat_cancers_vector <- estudio_vacunas_t_final_selected %>% 
  select(cancer_inmunoproliferativo, enfermedad_de_hodgkin, linfoma_no_hodgkin, leucemia) %>% 
  colnames(.)



estudio_vacunas_t_final_selected %<>% 
  mutate(
    variant_period = relevel(as.factor(variant_period), ref = "Dominance of Alpha variant"),
    across(comorb_no_cancer, ~if_else(.x == 1, "Sí", "No")),
    across(comorb_no_cancer, as.factor),
    across(comorb_no_cancer, ~relevel(.x, ref = "No")),
    across(c(hosp_stay_bin, icu_stay_bin), ~if_else(.x == 1, "Sí", "No")),
    across(c(hosp_stay_bin, icu_stay_bin), ~as.factor(.x)),
    across(c(hosp_stay_bin, icu_stay_bin), ~relevel(.x, ref = "No")),
  )
```

# Creating functions

```{r}
# metric performance
precision <- function(matrix) {
  # True positive
  tp <- matrix[2, 2]
  # false positive
  fp <- matrix[1, 2]
  return (tp / (tp + fp))
}

recall <- function(matrix) {
  # true positive
  tp <- matrix[2, 2]
  # false negative
  fn <- matrix[2, 1]
  return (tp / (tp + fn))
}

extract_coef_pval_OR <- function(model) {
  as.data.frame(tibble(
    term = names(model$coefficients),
    estimate = model$coefficients,
    std.error = sqrt(diag(vcov(model))),
    statistic = model$coefficients / sqrt(diag(vcov(model))),
    p.value = 2 * (1 - pnorm(abs(statistic))),
    OR = round(exp(estimate), 2),
    OR_CI_lower = round(exp(estimate - 1.96 * std.error), 2),
    OR_CI_upper = round(exp(estimate + 1.96 * std.error), 2)
  ))
}
```

# Logistic regression. Outcome: hosp_stay_bin

## Fitting the model

```{r}
logreg_hosp_stay_bin <- 
  glm(
    hosp_stay_bin ~ 
      age + desc_sexo + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc + 
      hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + 
      insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + 
      n_vacunas_covid + n_prev_infecc + ultima_inmun_infecc_bin,
    data =  estudio_vacunas_t_final_selected,
    family = binomial()
  )
```

## Adjusted Odds Ratios for hospital stay

```{r fig.height = 16, fig.width = 10}
coef_pval_df_hosp_stay_bin <- extract_coef_pval_OR(logreg_hosp_stay_bin)

# write.csv(
#   coef_pval_df_hosp_stay_bin,
#   "/opt/datos_compartidos/vacunas_data/tablas/coef_pval_df_hosp_stay_bin_manuscript.csv",
#   col.names = F
# )

htmlTable(coef_pval_df_hosp_stay_bin[,c("term","OR","OR_CI_lower","OR_CI_upper")])
```

## Forest Plot

```{r}
coef_pval_df_hosp_stay_bin2 <- 
  coef_pval_df_hosp_stay_bin %>% 
  mutate(
    orden = case_when(
      term == "age" ~ 1,
      term == "desc_sexoHombre" ~ 2,
      term == "asmaSí" ~ 3,
      term == "cardiopatia_isquemicaSí" ~ 4,
      term == "cirrosis_hepaticaSí" ~ 5,
      term == "diabetesSí" ~ 6,
      term == "epocSí" ~ 7,
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ 8,
      term == "hipertensionSí" ~ 9,
      term == "insuficiencia_cardiacaSí" ~ 10,
      term == "insuficiencia_renal_cronicaSí" ~ 11,
      term == "vihSí" ~ 12,
      term == "cancer_hematologicoSí" ~ 13,
      term == "cancer_organo_solidoSí" ~ 14,
      term == "n_vacunas_covidUna" ~ 15,
      term == "n_vacunas_covidDos" ~ 16,
      term == "n_vacunas_covidTres" ~ 17,
      term == "n_prev_infeccUno" ~ 18,
      term == "n_prev_infeccDos" ~ 19,
      term == "ultima_inmun_infecc_binOne to three months" ~ 20,
      term == "ultima_inmun_infecc_binThree to six months" ~ 21,
      term == "ultima_inmun_infecc_binSix months or more" ~ 22,
      term == "(Intercept)" ~ 23,
      TRUE ~ 999
    ),
    term = case_when(
      term == "age" ~ "Age (years)",
      term == "desc_sexoHombre" ~ "Sex (male)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "COPD",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic liver disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hipertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematological cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "n_vacunas_covidUna" ~ "1 vaccine dose",
      term == "n_vacunas_covidDos" ~ "2 vaccine doses",
      term == "n_vacunas_covidTres" ~ "3 vaccine doses",
      term == "n_prev_infeccUno" ~ "1 previous infection",
      term == "n_prev_infeccDos" ~ "2 previous infections",
      term == "ultima_inmun_infecc_binOne to three months" ~ "1-3 months past since last exposure*",
      term == "ultima_inmun_infecc_binThree to six months" ~ "3-6 months past since last exposure*",
      term == "ultima_inmun_infecc_binSix months or more" ~ "6+ months past since last exposure*",
      TRUE ~ term
    )
  ) %>% 
  arrange(desc(orden)) %>% 
  select(-orden) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(term = factor(term, levels = term, ordered = TRUE))


ggplot(
  coef_pval_df_hosp_stay_bin2, 
  aes(x = term, y = OR, ymin = OR_CI_lower, ymax = OR_CI_upper)) +
  geom_hline(yintercept = 1, colour = "blue", size = 0.5, linetype = "solid") +
  geom_pointrange(alpha = 0.5, shape = 18, size = 0.8) +
  geom_vline(xintercept = 0, size = 1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(
    breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8),
    expand = expansion(add = c(0.1, 0.1))
  ) +
  # scale_y_log10(
  #   limits=c(0.1, 4)
  #   ) +
  coord_flip() +
  labs(y = "Odds Ratio", x = NULL) +
  ggtitle("Adjusted Odds Ratios for hospital admission") +
  theme_minimal()

# Save the plot as a PNG file
ggsave(
  "/opt/datos_compartidos/vacunas_data/figuras/forest_plot_hosp_stay_bin_manuscript2.png",
  width = 3200,
  height = 1750,
  dpi = 300,
  units = "px"
)
```

## Model Performance

```{r eval = FALSE, include = FALSE}
y_test <- estudio_vacunas_final_selected$hosp_stay_bin

predict <- predict(logreg_hosp_stay_bin, estudio_vacunas_final_selected, type = 'response')
# confusion matrix
table_mat <- table(y_test, predict > 0.85)
table_mat

accuracy <- sum(diag(table_mat)) / sum(table_mat)
accuracy

prec <- precision(table_mat)
prec

rec <- recall(table_mat)
rec

f1 <- 2 * ((prec * rec) / (prec + rec))
f1

#AUTOC plot

predicted_bin<-as.numeric(c(predict > 0.1))
rocobj <- roc(y_test, predicted_bin)

auc <- round(auc(y_test, predicted_bin),2)

#create ROC plot
ggroc(rocobj, colour = 'steelblue', size = 2) +
  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')'))

metrics<-round(as.data.frame(rbind(accuracy,auc)),2)
metrics$metric<-rownames(metrics)
metrics$metric<-toupper(metrics$metric)

names(metrics)<-c('Value', 'Metric')

ggplot(data=metrics, aes(x=Metric, y=Value)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Value), vjust=1.6, color="white", size=3.5)+
  theme_minimal()
```

# Logistic regression. Outcome: icu_stay_bin

## Fitting the model

```{r}
logreg_icu_stay_bin <- 
  glm(
    icu_stay_bin ~ 
      age + desc_sexo + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc + 
      hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + 
      insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + 
      n_vacunas_covid + n_prev_infecc + ultima_inmun_infecc_bin,
    data = estudio_vacunas_t_final_selected,
    family = binomial()
  )

```

## Adjusted Odds ratio for ICU stay

```{r fig.height = 16, fig.width = 10}
coef_pval_df_logreg_icu_stay_bin <- extract_coef_pval_OR(logreg_icu_stay_bin)

# write.csv(
#   coef_pval_df_logreg_icu_stay_bin,
#   "/opt/datos_compartidos/vacunas_data/tablas/coef_pval_df_logreg_icu_stay_bin_manuscript.csv",
#   col.names = F
# )

htmlTable(coef_pval_df_logreg_icu_stay_bin[,c("term","OR","OR_CI_lower","OR_CI_upper")])
```

## Forest Plot

```{r}
coef_pval_df_logreg_icu_stay_bin2 <- 
  coef_pval_df_logreg_icu_stay_bin %>% 
  mutate(
    orden = case_when(
      term == "age" ~ 1,
      term == "desc_sexoHombre" ~ 2,
      term == "asmaSí" ~ 3,
      term == "cardiopatia_isquemicaSí" ~ 4,
      term == "cirrosis_hepaticaSí" ~ 5,
      term == "diabetesSí" ~ 6,
      term == "epocSí" ~ 7,
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ 8,
      term == "hipertensionSí" ~ 9,
      term == "insuficiencia_cardiacaSí" ~ 10,
      term == "insuficiencia_renal_cronicaSí" ~ 11,
      term == "vihSí" ~ 12,
      term == "cancer_hematologicoSí" ~ 13,
      term == "cancer_organo_solidoSí" ~ 14,
      term == "n_vacunas_covidUna" ~ 15,
      term == "n_vacunas_covidDos" ~ 16,
      term == "n_vacunas_covidTres" ~ 17,
      term == "n_prev_infeccUno" ~ 18,
      term == "n_prev_infeccDos" ~ 19,
      term == "ultima_inmun_infecc_binOne to three months" ~ 20,
      term == "ultima_inmun_infecc_binThree to six months" ~ 21,
      term == "ultima_inmun_infecc_binSix months or more" ~ 22,
      term == "(Intercept)" ~ 23,
      TRUE ~ 999
    ),
    term = case_when(
      term == "age" ~ "Age (years)",
      term == "desc_sexoHombre" ~ "Sex (male)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "COPD",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic liver disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hipertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematological cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "n_vacunas_covidUna" ~ "1 vaccine dose",
      term == "n_vacunas_covidDos" ~ "2 vaccine doses",
      term == "n_vacunas_covidTres" ~ "3 vaccine doses",
      term == "n_prev_infeccUno" ~ "1 previous infection",
      term == "n_prev_infeccDos" ~ "2 previous infections",
      term == "ultima_inmun_infecc_binOne to three months" ~ "1-3 months past since last exposure*",
      term == "ultima_inmun_infecc_binThree to six months" ~ "3-6 months past since last exposure*",
      term == "ultima_inmun_infecc_binSix months or more" ~ "6+ months past since last exposure*",
      TRUE ~ term
    )
  ) %>% 
  arrange(desc(orden)) %>% 
  select(-orden) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(
    term = factor(term, levels = term, ordered = TRUE),
    sobrepasa = case_when(
      OR_CI_upper > (sort(coef_pval_df_logreg_icu_stay_bin$OR_CI_upper, decreasing = TRUE)[2] + 10) ~ 1,
      TRUE ~ 0
    ),
    nuevo_OR_CI_upper = case_when(
      OR_CI_upper > (sort(coef_pval_df_logreg_icu_stay_bin$OR_CI_upper, decreasing = TRUE)[2] + 10)
        ~ sort(coef_pval_df_logreg_icu_stay_bin$OR_CI_upper, decreasing = TRUE)[2] + 0.1,
      TRUE ~ OR_CI_upper
    )
  )



ggplot(
  coef_pval_df_logreg_icu_stay_bin2, 
  aes(x = term, y = OR, ymin = OR_CI_lower, ymax = nuevo_OR_CI_upper)) +
  geom_hline(yintercept = 1, colour = "blue", size = 0.5, linetype = "solid") +
  geom_pointrange(alpha = 0.5, shape = 18, size = 0.8) +
  geom_vline(xintercept = 0, size = 1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  geom_text(
    data = coef_pval_df_logreg_icu_stay_bin2 %>% filter(sobrepasa == 1),
    mapping = aes(x = term, y = nuevo_OR_CI_upper),
    label = "▶", size = 3, family = "HiraKakuPro-W3", colour = "red", vjust = 0.4) +
  # geom_point(data = coef_pval_df_logreg_icu_stay_bin2 %>% filter(sobrepasa = 1),
  #            shape = 17,
  #            ) +
  scale_y_continuous(expand = expansion(add = c(0.1, 0.05))) +
  # scale_y_log10(limits=c(0.1, 4)) +
  coord_flip() +
  labs(y = "Odds Ratio", x = NULL) +
  ggtitle("Adjusted Odds Ratios for ICU admission") +
  theme_minimal()

# Save the plot as a PNG file
ggsave(
  "/opt/datos_compartidos/vacunas_data/figuras/forest_plot_icu_stay_bin_manuscript2.png",
  width = 3200,
  height = 1750,
  dpi = 300,
  units = "px"
)
```

## Model Performance

```{r eval = FALSE, include = FALSE}
y_test<-estudio_vacunas_final_selected$icu_stay_bin

predict <- predict(logreg_icu_stay_bin, estudio_vacunas_final_selected, type = 'response')
# confusion matrix
table_mat <- table(y_test, predict > 0.1)
table_mat

accuracy <- sum(diag(table_mat)) / sum(table_mat)
accuracy

prec <- precision(table_mat)
prec

rec <- recall(table_mat)
rec

f1 <- 2 * ((prec * rec) / (prec + rec))
f1

#AUTOC plot

predicted_bin<-as.numeric(c(predict > 0.3))
rocobj <- roc(y_test, predicted_bin)

auc <- round(auc(y_test, predicted_bin),2)

#create ROC plot
ggroc(rocobj, colour = 'steelblue', size = 2) +
  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')'))

metrics<-round(as.data.frame(rbind(accuracy,auc)),2)
metrics$metric<-rownames(metrics)
metrics$metric<-toupper(metrics$metric)

names(metrics)<-c('Value', 'Metric')

ggplot(data=metrics, aes(x=Metric, y=Value)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Value), vjust=1.6, color="white", size=3.5)+
  theme_minimal()
```

# Logistic regression. Outcome: mortality30d

## Fitting the model

```{r}
logreg_mortality <- 
  glm(
    mortality30d ~ 
      age + desc_sexo + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc + 
      hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + 
      insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + 
      n_vacunas_covid + n_prev_infecc + ultima_inmun_infecc_bin,
    data = estudio_vacunas_t_final_selected,
    family = binomial()
  )

```

## Adjusted Odds ratios for mortality within 30 days past infection

```{r fig.height = 16, fig.width = 10}
coef_pval_df_logreg_mortality <- extract_coef_pval_OR(logreg_mortality)

# write.csv(
#   coef_pval_df_logreg_mortality,
#   "/opt/datos_compartidos/vacunas_data/tablas/coef_pval_df_logreg_mortality_manuscript.csv",
#   col.names = F
# )

htmlTable(coef_pval_df_logreg_mortality[,c("term","OR","OR_CI_lower","OR_CI_upper")])
```

## Forest Plot

```{r}
coef_pval_df_logreg_mortality2 <- 
  coef_pval_df_logreg_mortality %>% 
  mutate(orden = case_when(
    term == "age" ~ 1,
      term == "desc_sexoHombre" ~ 2,
      term == "asmaSí" ~ 3,
      term == "cardiopatia_isquemicaSí" ~ 4,
      term == "cirrosis_hepaticaSí" ~ 5,
      term == "diabetesSí" ~ 6,
      term == "epocSí" ~ 7,
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ 8,
      term == "hipertensionSí" ~ 9,
      term == "insuficiencia_cardiacaSí" ~ 10,
      term == "insuficiencia_renal_cronicaSí" ~ 11,
      term == "vihSí" ~ 12,
      term == "cancer_hematologicoSí" ~ 13,
      term == "cancer_organo_solidoSí" ~ 14,
      term == "n_vacunas_covidUna" ~ 15,
      term == "n_vacunas_covidDos" ~ 16,
      term == "n_vacunas_covidTres" ~ 17,
      term == "n_prev_infeccUno" ~ 18,
      term == "n_prev_infeccDos" ~ 19,
      term == "ultima_inmun_infecc_binOne to three months" ~ 20,
      term == "ultima_inmun_infecc_binThree to six months" ~ 21,
      term == "ultima_inmun_infecc_binSix months or more" ~ 22,
      term == "(Intercept)" ~ 23,
      TRUE ~ 999
    ),
    term = case_when(
      term == "age" ~ "Age (years)",
      term == "desc_sexoHombre" ~ "Sex (male)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "COPD",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic liver disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hipertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematological cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "n_vacunas_covidUna" ~ "1 vaccine dose",
      term == "n_vacunas_covidDos" ~ "2 vaccine doses",
      term == "n_vacunas_covidTres" ~ "3 vaccine doses",
      term == "n_prev_infeccUno" ~ "1 previous infection",
      term == "n_prev_infeccDos" ~ "2 previous infections",
      term == "ultima_inmun_infecc_binOne to three months" ~ "1-3 months past since last exposure*",
      term == "ultima_inmun_infecc_binThree to six months" ~ "3-6 months past since last exposure*",
      term == "ultima_inmun_infecc_binSix months or more" ~ "6+ months past since last exposure*",
      TRUE ~ term)) %>% 
  arrange(desc(orden)) %>% 
  select(-orden) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(term = factor(term, levels = term, ordered = TRUE))


ggplot(
  coef_pval_df_logreg_mortality2, 
  aes(x = term, y = OR, ymin = OR_CI_lower, ymax = OR_CI_upper)) +
  geom_hline(yintercept = 1, colour = "blue", size = 0.5, linetype = "solid") +
  geom_pointrange(alpha = 0.5, shape = 18, size = 0.8) +
  geom_vline(xintercept = 0, size = 1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  # scale_y_log10(limits=c(0.1, 4)) +
  coord_flip() +
  labs(y = "Odds Ratio", x = NULL) +
  ggtitle("Adjusted Odds Ratio for mortality within 30 days from infection") +
  theme_minimal()



ggsave(
  "/opt/datos_compartidos/vacunas_data/figuras/forest_plot_mortality30d_manuscript2.png",
  width = 3200,
  height = 1750,
  dpi = 300,
  units = "px"
)
```

# Data subsetting prior to regressions

```{r}
# A function created to make my life a tiny bit easier. 
# Not fancy, but practical
subsetting <- function(lev_vacc_vctr, age_gr, variant) {
  output <- estudio_vacunas_t_final_selected %>% 
    filter(
      n_vacunas_covid %in% lev_vacc_vctr
      & age_fct_60years == age_gr
      & variant_period == variant
    ) %>% 
    mutate(across(c(n_vacunas_covid, age_fct_60years, variant_period), droplevels))
  return(output)
}

# Subsetting: two doses
twodoses_over60alpha <- subsetting(c("Dos", "Ninguna"), "60 años o más", "Dominance of Alpha variant")
twodoses_over60delta <- subsetting(c("Dos", "Ninguna"), "60 años o más", "Dominance of Delta variant")
twodoses_over60omicron <- subsetting(c("Dos", "Ninguna"), "60 años o más", "Dominance of Omicron variant")

twodoses_under60alpha <- subsetting(c("Dos", "Ninguna"), "Menor de 60 años", "Dominance of Alpha variant")
twodoses_under60delta <- subsetting(c("Dos", "Ninguna"), "Menor de 60 años",  "Dominance of Delta variant")
twodoses_under60omicron <- subsetting(c("Dos", "Ninguna"), "Menor de 60 años", "Dominance of Omicron variant")

# Subsetting: three doses
threedoses_over60alpha <- subsetting(c("Tres", "Ninguna"), "60 años o más", "Dominance of Alpha variant")
threedoses_over60delta <- subsetting(c("Tres", "Ninguna"), "60 años o más", "Dominance of Delta variant")
threedoses_over60omicron <- subsetting(c("Tres", "Ninguna"), "60 años o más", "Dominance of Omicron variant")

threedoses_under60alpha <- subsetting(c("Tres", "Ninguna"), "Menor de 60 años", "Dominance of Alpha variant")
threedoses_under60delta <- subsetting(c("Tres", "Ninguna"), "Menor de 60 años",  "Dominance of Delta variant")
threedoses_under60omicron <- subsetting(c("Tres", "Ninguna"), "Menor de 60 años", "Dominance of Omicron variant")
```

# Logistic regressions adjusted for age group and variant. Outcome: HOSPITAL ADMISSION

```{r}
flights_recipe <- 
  recipe(
    arr_delay ~ ., 
    # Model outcome should be placed in LHS. predictors should be found in RHS.
    # Variables may be listed by name, or you can use the dot (.) to indicate all other variables
    data = flight_data
    # This doesn’t change the data itself; it is only used as a reference for the variables...
    # and their types, like factors, integers, dates, etc
  ) %>% 
  update_role(flight, time_hour, new_role = "ID") %>% 
  # new role "ID" is custom; it is made as we name it inside the function
  # This tells the recipe to keep the two variables as NEITHER predictor NOR outcome (out of formula)
  step_dummy(all_nominal_predictors()) %>% 
  # For factor variables, standard practice is to convert them into dummy or indicator variables...
  # to make them numeric. These are binary values for each level of the factor.
  # Unlike the standard model formula methods in R, a recipe does not automatically create...
  # dummy variables for you; you’ll need to tell your recipe to add this step (step_dummy)
  step_zv(all_predictors())
  # Since factor levels and thus dummy variables come from reference data, when the recipe is applied...
  # to datasets that do not contain some levels in a variable, the correspondant dummy column...
  # would contain all zeros. This is a “zero-variance predictor” that has no information...
  # within the column. It usually causes warnings and other issues. step_zv() will remove columns...
  # from the data when the variable have a single value.


colnames(estudio_vacunas_t_final_selected)
vars_to_exclude <- estudio_vacunas_t_final_selected %>% 
  select() %>% 
  colnames(.)

custom_recipe <- function(model_outcome, data) {
  recipe <- 
    recipe({{model_outcome}} ~ ., data = data) %>% 
    update_role(flight, time_hour, new_role = "ID")
}






linear_regr_fit <- 
  workflow() %>% 
  add_model(logistic_reg()) %>% 
  add_recipe(custom_recipe(model_outcome = , data = .....)) %>% 
  fit(data = .....)









vars_to_exclude <- estudio_vacunas_t_final_selected %>% 
  select(
    nuhsa_encriptado:prev_infect_date, fec_fallecimiento, mortality,
    cancer_colorrectal:cancer_inmunoproliferativo, enfermedad_de_hodgkin, leucemia:sarcoma_de_kaposi,
    cod_fec_ingreso:estancia_hospitalaria_dias, infeccion_ingreso_dias, infeccion_alta_dias,
    ultima_inmun_infecc_meses_quant
    
  ) %>% 
  colnames(.)

or_glm_under60_alpha_2doses <- 
  glm(
    hosp_stay_bin ~ 
      age + desc_sexo + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc + 
      hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + 
      insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + 
      n_vacunas_covid + n_prev_infecc,
    data = under60_alpha_2doses,
    family = binomial()
  ) %>% 
  extract_coef_pval_OR() %>% 
  mutate(
    VE = (1 - OR) * 100,
    VE_CI_lower = (1 - OR_CI_lower) * 100,
    VE_CI_upper = (1 - OR_CI_upper) * 100,
    age_group = "Under 60 years old",
    variant = "Alpha",
    dose_comparison = "2 doses"
  )




summary(over60_alpha_2doses)
over60_alpha_2doses <- estudio_vacunas_t_final_selected %>% 
  filter(
    age_fct_60years == "60 años o más"
    & variant_period == "Dominance of Alpha variant"
    & n_vacunas_covid %in% c("Dos", "Ninguna")
  ) %>% 
  mutate(n_vacunas_covid = droplevels(n_vacunas_covid))
sapply(lapply(over60_alpha_2doses, unique), length)
or_glm_over60_alpha_2doses <- 
  glm(
    hosp_stay_bin ~ 
      age + 
      desc_sexo + 
      asma + 
      cardiopatia_isquemica + 
      cirrosis_hepatica + 
      diabetes + 
      epoc + 
      hepatopatia_cronica_excepto_cirrosis + 
      hipertension + 
      insuficiencia_cardiaca + 
      insuficiencia_renal_cronica +
      # vih +
      cancer_hematologico +
      cancer_organo_solido + 
      n_vacunas_covid +
      n_prev_infecc +
      ultima_inmun_infecc_bin
    ,
    data = over60_alpha_2doses,
    family = binomial()
  ) %>% 
  extract_coef_pval_OR() %>% 
  mutate(
    VE = (1 - OR) * 100,
    VE_CI_lower = (1 - OR_CI_lower) * 100,
    VE_OR_CI_upper = (1 - OR_CI_upper) * 100,
    age_group = "60 years old or more",
    variant = "Alpha",
    dose_comparison = "2 doses"
  )

# para dos dosis incluir desde los 12 años. para tres dosis, incluir desde los 18 años
```

