---
title: "Figures for the Andalusia COVID vaccines paper"
author: "Á. Serrano-Ortiz & JL Romero-Cabrera; EPIDATA team"
output: 
  html_document:
    theme: lumen
---



```{r setup, include = FALSE}
library(dplyr)
library(magrittr)
library(ggplot2)
library(janitor)
library(stringi)
library(stringr)
library(lubridate)
library(rio)
library(readr)
library(kableExtra)
library(networkD3)
```

```{r, include = FALSE}
##################################################################################
####################### FUNCIONES Y VALORES PARAMETRIZADOS #######################
##################################################################################



# Función para leer archivos ""CSV"" (excel siempre dando problemas) y asignar nombres
# hashtag grasiaschatgptporenseñarnosR

leer_csv_carpeta <- function(ruta) {
  rutas_archivos <- list.files(path = ruta, pattern = ".xls", full.names = TRUE)
  lista <- list()
  for (i in seq_along(rutas_archivos)) {
    datos <- read_csv(rutas_archivos[i])
    nombre <- str_remove(rutas_archivos[i], fixed(ruta))
    datos <- datos %>% 
      mutate(
        outcome = if_else(
          grepl("intensivecare_only_hosp", nombre),
          "intensivecare_only_hosp",
          str_extract(nombre, "(?<=_)[^_]+")
        ),
        veffec = (1 - OR)*100,
        veffec_lower = (1 - OR_lower)*100,
        veffec_upper = (1 - OR_upper)*100
      ) %>% 
      select(outcome, everything())
    lista[[nombre]] <- datos
  }
  return(lista)
}



colours <- c("#FFD92F", "#8DA0CB", "#A6D854", "#E78AC3")
colours_vaccines <- c("#D3BA68", "#D5695D", "#5D8CA8")
```

## Variants distribution in Spain {.tabset .tabset-fade .tabset-pills}

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
ecdc_variants_history <- import("https://opendata.ecdc.europa.eu/covid19/virusvariant/csv/data.csv") %>% 
  filter(source == "GISAID" & country == "Spain")



ecdc_all_variants <- ecdc_variants_history %>% 
  pull(variant) %>% 
  unique(.)

countA <- ecdc_variants_history %>% 
  count(variant, name = "count") %>% 
  slice_head(n = ceiling(nrow(.)/2))

countB <- ecdc_variants_history %>% 
  count(variant, name = "count") %>% 
  slice_tail(n = floor(nrow(.)/2))

table_variants_spain <- kable(list(countA, countB)) %>% kable_styling()



variants_who_labels <- tribble(
  ~who_label_family, ~lineage_and_additional_mutations,
  "Omicron", "BA.2.75",
  "Omicron", "BQ.1",
  "Omicron", "XBB",
  "Omicron", "XBB.1.5",
  "Omicron", "CH.1.1",
  "Omicron", "XBB.1.16",
  "Omicron", "FE.1",
  "Alpha", "B.1.1.7",
  "Alpha", "B.1.1.7+E484K",
  "Epsilon", "B.1.427/B.1.429",
  "Unnamed", "B.1.616",
  "Eta", "B.1.525",
  "Theta", "P.3",
  "Kappa", "B.1.617.1",
  "Unnamed", "B.1.620",
  "Unnamed", "B.1.617.3",
  "Unnamed", "B.1.214.2",
  "Unnamed", "A.23.1+E484K",
  "Unnamed", "A.27",
  "Unnamed", "A.28",
  "Unnamed", "C.16",
  "Beta", "B.1.351+P384L",
  "Beta", "B.1.351+E516Q",
  "Alpha", "B.1.1.7+L452R",
  "Alpha", "B.1.1.7+S494P",
  "Iota", "B.1.526",
  "Unnamed", "B.1.526.1",
  "Unnamed", "B.1.526.2",
  "Zeta", "P.2",
  "Unnamed", "B.1.1.519",
  "Unnamed", "AV.1",
  "Unnamed", "AT.1",
  "Unnamed", "C.36+L452R",
  "Gamma", "P.1+P681H",
  "Mu", "B.1.621",
  "Lambda", "C.37",
  "Unnamed", "AY.4.2",
  "Unnamed", "B.1.1.318",
  "Delta", "B.1.617.2",
  "Unnamed", "C.1.2",
  "Delta", "B.1.617.2",
  "Delta", "B.1.617.2",
  "Delta", "B.1.617.2",
  "Beta", "B.1.351",
  "Gamma", "P.1",
  "Unnamed", "B.1.640",
  "Unnamed", "XF",
  "Unnamed", "XD",
  "Delta", "B.1.617.2",
  "Omicron", "BA.1",
  "Omicron", "BA.3",
  "Omicron", "BA.2 + L452X",
  "Omicron", "XAK",
  "Omicron", "B.1.1.529 + R346X",
  "Omicron", "B.1.1.529 + K444X, N460X",
  "Omicron", "B.1.1.529 + N460X, F490X",
  "Omicron", "BA.2.3.20",
  "Omicron", "BF.7",
  "Omicron", "BA.2",
  "Omicron", "BA.4",
  "Omicron", "BA.5",
  "Omicron", "XBC",
  "Omicron", "BN.1",
  "Omicron", "XAY"
) %>% 
  mutate(
    lineage_and_additional_mutations = if_else(
      who_label_family == "Unnamed", "Unknown", lineage_and_additional_mutations
    )
  ) %>% 
  distinct()



variants_week_spain_labels <- ecdc_variants_history %>% 
  select(-c(country, country_code, source, valid_denominator)) %>%
  mutate(variant = if_else(variant %in% c("Other", "UNK"), "Unknown", variant)) %>% 
  left_join(variants_who_labels, by = c("variant" = "lineage_and_additional_mutations")) %>% 
  mutate(
    who_label_family = case_when(
      who_label_family %in% c("Omicron", "Delta", "Alpha", "Gamma", "Beta") ~ who_label_family,
      TRUE ~ "Other variants or unnamed"
    ),
    percent_variant2 = number_detections_variant/number_sequenced*100
  ) %>% 
  filter(!is.nan(percent_variant2)) %>% 
  mutate(year_week2 = gsub("-", "", year_week)) %>% 
  filter(
    !year_week2 %in% c(
      "202006", paste0(2022, 14:52), paste0(20230, 1:9), paste0(2023, 10:23)
    )) %>% 
  select(who_label_family, percent_variant2, everything())



variants_totalpercents_ranking <- variants_week_spain_labels %>% 
  group_by(who_label_family) %>% 
  summarise(weektotal_percent_sum = sum(percent_variant2)) %>% 
  arrange(desc(weektotal_percent_sum))

variants_ordered <- c("Omicron", "Delta", "Alpha", "Beta", "Gamma", "Other variants or unnamed")

variants_week_spain_labels %<>% 
  mutate(who_label_family = factor(who_label_family, levels = variants_ordered, ordered = TRUE))
  


cutpoints_variants_of_study <- variants_week_spain_labels %>% 
  filter(who_label_family %in% c("Alpha", "Delta", "Omicron")) %>% 
  mutate(
    year_week = gsub("-", "", year_week),
    year_week = as.numeric(year_week)
  ) %>% 
  arrange(who_label_family, year_week)



begining_alpha <- 
  ymd(20210308) # monday of isoweek 2021-10, when alpha started to account for >=75% of sequenced samples
end_alpha <- 
  ymd(20210606)
begining_delta <- 
  ymd(20210712)
end_delta <- 
  ymd(20211212)
begining_omicron <- 
  ymd(20211227)
end_omicron <- 
  ymd(20220331)



variants_week_spain_labels2 <- variants_week_spain_labels %>% 
  group_by(who_label_family, year_week) %>% 
  summarise(percent_variant2 = sum(percent_variant2)) %>% 
  ungroup()
  
alpha_col <- "#CBD5E8"
delta_col <- "#FDCDAC"
omicron_col <- "#B3E2CD"
```

### Option A

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
variants_week_spain_labels2 %>% 
  ggplot() +
  geom_rect(
    xmin = "2021-10", xmax = "2021-22",
    ymin = 0, ymax = 100,
    fill = alpha_col, colour = NA, alpha = 0.2) +
  geom_rect(
    xmin = "2021-28", xmax = "2021-49",
    ymin = 0, ymax = 100,
    fill = delta_col, colour = NA, alpha = 0.2) +
  geom_rect(
    xmin = "2021-52", xmax = "2022-13",
    ymin = 0, ymax = 100,
    fill = omicron_col, colour = NA, alpha = 0.2) +
  geom_line(
    mapping = aes(x = year_week, y = percent_variant2, group = who_label_family, colour = who_label_family),
    size = 1.5,
    alpha = 1
  ) +
  scale_y_continuous(expand = expansion(add = c(0, 1))) +
  # scale_x_date(expand = expansion(add = c(days(0), days(0)))) +
  # scale_colour_brewer(palette = "Set2") +
  scale_colour_manual(values = rev(c("#FFD92F", "#A6D854", "#E78AC3", "#8DA0CB", "#FC8D62", "#66C2A5"))) +
  theme_minimal() + 
  theme(
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 90, size = 6, hjust = 1, vjust = 0.5),
    legend.title = element_blank()
  ) +
  labs(
    x = "Epi week",
    y = "Variants percent",
    title = "Weekly proportion of SARS-CoV-2 variants sequenced in Spain",
    caption = "Source: GISAID, ECDC"
  )
```

### Option B

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
variants_week_spain_labels2 %>% 
  ggplot() +
  geom_col(
    mapping = aes(x = year_week, y = percent_variant2, fill = who_label_family),
    colour = NA,
    width = 1,
    alpha = 0.7
  ) +
  # Alpha variant label and line
  # geom_segment(
  #   x = "2021-10", xend = "2021-22",
  #   y = 20, yend = 20,
  #   colour = "red", size = 4, linetype = "solid",
  #   arrow = arrow(length = unit(4, "mm"))
  # ) +
  geom_vline(xintercept = "2021-10", colour = "red", size = 0.5, linetype = "solid") +
  geom_text(
    x = "2021-10", y = 20, label = "Alpha variant in 75% of sequenced samples",
    size = 3, colour = "red", angle = 90, vjust = 1.2, hjust = 0.3
  ) +
  # Delta variant label and line
  geom_vline(xintercept = "2021-28", colour = "grey25", size = 0.5, linetype = "solid") +
  geom_text(
    x = "2021-28", y = 20, label = "Delta variant at least 75%",
    size = 3, colour = "grey25", angle = 90, vjust = 1.2, hjust = 0.3
  ) +
  # Omicron variant label and line
  geom_vline(xintercept = "2021-52", colour = "grey25", size = 0.5, linetype = "solid") +
  geom_text(
    x = "2021-52", y = 20, label = "Omicron variant at least 75%",
    size = 3, colour = "grey25", angle = 90, vjust = 1.2, hjust = 0.3
  ) +
  scale_y_continuous(expand = expansion(add = c(0, 0))) +
  scale_x_discrete(expand = expansion(add = c(days(0), days(0)))) +
  scale_fill_brewer(palette = "Set3") +
  theme_bw() + 
  theme(
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 90, size = 6, hjust = 1, vjust = 0.5),
    legend.title = element_blank()
  ) +
  labs(
    x = "Epi week",
    y = "Variants percent",
    title = "Weekly proportion of SARS-CoV-2 variants sequenced in Spain",
    caption = "Source: GISAID, ECDC"
  )



# ggsave(
#   filename = "/opt/datos_compartidos/vacunas_data/figuras/variants_spain.png",
#   device = "png",
#   width = 8000,
#   height = 8000,
#   dpi = 700,
#   units = "px"
# )
```

## Plots for vaccine effectiveness sub-analyses {.tabset .tabset-fade .tabset-pills}

```{r, include = FALSE}
# CARGA DE DATOS
# 
# Para el/la que quiera consultar el código: el nombre de los objetos obedece a
# 
# - en **minúscula**, los criterios para subdividir la base de datos;
# - en **mayúscula**, el outcome principal incluido en las regresiones GAM empleadas por cada subgrupo. 
# Son mutuamente excluyentes (si se incluyó el n_vacunas, se excluyó el tiempo desde inmunización y viceversa)



var_agegr_VACCINES <-
  leer_csv_carpeta(
    ruta = "/opt/datos_compartidos/vacunas_data/tablas/groups by variant & agegroup (logreg nvaccines only)"
  ) %>%
  bind_rows() %>%
  mutate(
    agegroup = str_extract(original_model, "(?<=_)[^_]+$"),
    agegroup = if_else(agegroup == "60", "60 years or older", "59 years or younger"),
    variant = str_extract(original_model, "^[^_]+"),
    variant = if_else(variant == "Before", "Before α", variant),
    predictor = case_when(
      predictor == "n_vacunas_covidUna" ~ "One",
      predictor == "n_vacunas_covidDos" ~ "Two",
      predictor == "n_vacunas_covidTres" ~ "Three"
    ),
    predictor = factor(predictor, levels = c("One", "Two", "Three"))
  ) %>%
  select(outcome, predictor, variant, agegroup, everything())



var_agegr_LASTIMMUN <-
  leer_csv_carpeta(
    ruta = "/opt/datos_compartidos/vacunas_data/tablas/groups2 by variant & agegroup (logreg lastimmun only)"
  ) %>%
  bind_rows() %>%
  mutate(
    agegroup = str_extract(original_model, "(?<=_)[^_]+$"),
    agegroup = if_else(agegroup == "60", "60 years or older", "59 years or younger"),
    variant = str_extract(original_model, "^[^_]+"),
    variant = if_else(variant == "Before", "Before α", variant),
    predictor = case_when(
      predictor == "ultima_inmun_infecc_binNone" ~ "None",
      predictor == "ultima_inmun_infecc_bin31-60" ~ "31-60",
      predictor == "ultima_inmun_infecc_bin61-90" ~ "61-90",
      predictor == "ultima_inmun_infecc_bin91-120" ~ "91-120",
      predictor == "ultima_inmun_infecc_bin121-180" ~ "121-180",
      predictor == "ultima_inmun_infecc_bin>180" ~ ">180"
    ),
    predictor = factor(
      predictor,
      levels = c("None", "31-60", "61-90", "91-120", "121-180", ">180"))
  ) %>%
  select(outcome, predictor, variant, agegroup, everything())



vacc_agegroup_LASTIMMUN <-
  leer_csv_carpeta(
    ruta = "/opt/datos_compartidos/vacunas_data/tablas/groups2 by nvaccines & agegroup (logreg lastimmun only)"
  ) %>%
  bind_rows() %>%
  mutate(
    nvaccine = str_extract(original_model, "(?<=_)[^_]+$"),
    nvaccine = case_when(
      nvaccine == "Ninguna" ~ "None", 
      nvaccine == "Una" ~ "1 dose", 
      nvaccine == "Dos" ~ "2 doses", 
      nvaccine == "Tres" ~ "3 doses"
    ),
    nvaccine = factor(nvaccine, levels = c("None", "1 dose", "2 doses", "3 doses")),
    agegroup = str_extract(original_model, "^[^_]+"),
    agegroup = if_else(agegroup == "60", "60 years or older", "59 years or younger"),
    predictor = case_when(
      predictor == "ultima_inmun_infecc_binNone" ~ "None",
      predictor == "ultima_inmun_infecc_bin31-60" ~ "31-60",
      predictor == "ultima_inmun_infecc_bin61-90" ~ "61-90",
      predictor == "ultima_inmun_infecc_bin91-120" ~ "91-120",
      predictor == "ultima_inmun_infecc_bin121-180" ~ "121-180",
      predictor == "ultima_inmun_infecc_bin>180" ~ ">180"
    ),
    predictor = factor(
      predictor,
      levels = c("None", "31-60", "61-90", "91-120", "121-180", ">180"))
  ) %>%
  select(outcome, predictor, nvaccine, agegroup, everything())
```

### SUB-A1: Vaccine effectiveness by number of vaccine doses (segregated by variant period and age group) 

Groups manually excluded:

- Cases before alpha rise, 60 years old or above, with more than one vaccine dose.
- All cases before alpha rise, younger than 60 years.
- Cases during alpha predominance, 60 years old or above, with three vaccine doses.
- All cases during alpha predominance, younger than 60 years.
- Cases during delta predominance, younger than 60 years, with three vaccine doses.

Groups automatically excluded:

- Those with upper vaccine effectiveness estimation interval above 100%.
- Those with lower vaccine effectiveness estimation interval below 0%.
- Those with vaccine effectiveness point-estimate above 100% or below 0%.


```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
# INITIAL FILTERS WERE INSERTED MANUALLY AFTER DISCUSSION BETWEEN MAIN AUTHORS

var_agegr_VACCINES %>% 
  filter(outcome == "hospitalisation") %>% 
  filter(
    (original_model == "Before_60" & predictor == "One")
    | (original_model == "α_60" & predictor %in% c("One", "Two"))
    | (original_model == "Δ_Menor" & predictor %in% c("One", "Two"))
    | (original_model == "Δ_60" & predictor %in% c("One", "Two", "Three"))
    | original_model %in% c("Ό_60", "Ό_Menor")
  ) %>%
  filter(veffec_lower > 0 & veffec_lower <= 100) %>% 
  filter(veffec_upper > 0 & veffec_upper <= 100) %>%
  filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = veffec_lower, ymax = veffec_upper, colour = variant),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = veffec, colour = variant), size = 3) +
  facet_wrap(~agegroup, nrow = 2, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(
    y = "Vaccine effectiveness",
    x = "Vaccine doses",
    colour = "Variant period",
    title = "Outcome: hospitalisation"
  )



var_agegr_VACCINES %>% 
  filter(outcome == "intensivecare") %>% 
  filter(
    (original_model == "Before_60" & predictor == "One")
    | (original_model == "α_60" & predictor %in% c("One", "Two"))
    | (original_model == "Δ_Menor" & predictor %in% c("One", "Two"))
    | (original_model == "Δ_60" & predictor %in% c("One", "Two", "Three"))
    | original_model %in% c("Ό_60", "Ό_Menor")
  ) %>%
  filter(veffec_lower > 0 & veffec_lower <= 100) %>% 
  filter(veffec_upper > 0 & veffec_upper <= 100) %>%
  filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = veffec_lower, ymax = veffec_upper, colour = variant),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = veffec, colour = variant), size = 3) +
  facet_wrap(~agegroup, nrow = 2, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(
    y = "Vaccine effectiveness",
    x = "Vaccine doses",
    colour = "Variant period",
    title = "Outcome: ICU admission"
  )



var_agegr_VACCINES %>%
  filter(outcome == "mortality") %>% 
  filter(
  (original_model == "Before_60" & predictor == "One")
  | (original_model == "α_60" & predictor %in% c("One", "Two"))
  | (original_model == "Δ_Menor" & predictor %in% c("One", "Two"))
  | (original_model == "Δ_60" & predictor %in% c("One", "Two", "Three"))
  | original_model %in% c("Ό_60", "Ό_Menor")
  ) %>%
  filter(veffec_lower > 0 & veffec_lower <= 100) %>% 
  filter(veffec_upper > 0 & veffec_upper <= 100) %>%
  filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = veffec_lower, ymax = veffec_upper, colour = variant),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = veffec, colour = variant), size = 3) +
  facet_wrap(~agegroup, nrow = 2, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(
    y = "Vaccine effectiveness",
    x = "Vaccine doses",
    colour = "Variant period",
    title = "Outcome: 30-days mortality"
  )
```

### SUB-A2: NO MANUAL FILTERS Vaccine effectiveness by number of vaccine doses (segregated by variant period and age group) 

Groups automatically excluded:

- Those with upper vaccine effectiveness estimation interval above 100%.
- Those with lower vaccine effectiveness estimation interval below 0%.
- Those with vaccine effectiveness point-estimate above 100% or below 0%.


```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
# INITIAL FILTERS WERE INSERTED MANUALLY AFTER DISCUSSION BETWEEN MAIN AUTHORS

var_agegr_VACCINES %>% 
  filter(outcome == "hospitalisation") %>% 
  # filter(
  #   (original_model == "Before_60" & predictor == "One")
  #   | (original_model == "α_60" & predictor %in% c("One", "Two"))
  #   | (original_model == "Δ_Menor" & predictor %in% c("One", "Two"))
  #   | (original_model == "Δ_60" & predictor %in% c("One", "Two", "Three"))
  #   | original_model %in% c("Ό_60", "Ό_Menor")
  # ) %>%
  filter(veffec_lower > 0 & veffec_lower <= 100) %>% 
  filter(veffec_upper > 0 & veffec_upper <= 100) %>%
  filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = veffec_lower, ymax = veffec_upper, colour = variant),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = veffec, colour = variant), size = 3) +
  facet_wrap(~agegroup, nrow = 2, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(
    y = "Vaccine effectiveness",
    x = "Vaccine doses",
    colour = "Variant period",
    title = "Outcome: hospitalisation"
  )



var_agegr_VACCINES %>% 
  filter(outcome == "intensivecare") %>% 
  # filter(
  #   (original_model == "Before_60" & predictor == "One")
  #   | (original_model == "α_60" & predictor %in% c("One", "Two"))
  #   | (original_model == "Δ_Menor" & predictor %in% c("One", "Two"))
  #   | (original_model == "Δ_60" & predictor %in% c("One", "Two", "Three"))
  #   | original_model %in% c("Ό_60", "Ό_Menor")
  # ) %>%
  filter(veffec_lower > 0 & veffec_lower <= 100) %>% 
  filter(veffec_upper > 0 & veffec_upper <= 100) %>%
  filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = veffec_lower, ymax = veffec_upper, colour = variant),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = veffec, colour = variant), size = 3) +
  facet_wrap(~agegroup, nrow = 2, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(
    y = "Vaccine effectiveness",
    x = "Vaccine doses",
    colour = "Variant period",
    title = "Outcome: ICU admission"
  )



var_agegr_VACCINES %>%
  filter(outcome == "mortality") %>% 
  # filter(
  # (original_model == "Before_60" & predictor == "One")
  # | (original_model == "α_60" & predictor %in% c("One", "Two"))
  # | (original_model == "Δ_Menor" & predictor %in% c("One", "Two"))
  # | (original_model == "Δ_60" & predictor %in% c("One", "Two", "Three"))
  # | original_model %in% c("Ό_60", "Ό_Menor")
  # ) %>%
  filter(veffec_lower > 0 & veffec_lower <= 100) %>% 
  filter(veffec_upper > 0 & veffec_upper <= 100) %>%
  filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = veffec_lower, ymax = veffec_upper, colour = variant),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = veffec, colour = variant), size = 3) +
  facet_wrap(~agegroup, nrow = 2, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(
    y = "Vaccine effectiveness",
    x = "Vaccine doses",
    colour = "Variant period",
    title = "Outcome: 30-days mortality"
  )
```

### SUB-B1: Vaccine effectiveness by time since last exposure: infection/vaccination (segregated by variant period and age group)

Groups manually excluded:

- All cases before alpha rise, younger than 60 years.
- All cases during alpha predominance, younger than 60 years.

Groups automatically excluded:

- Those with upper vaccine effectiveness estimation interval above 100%.
- Those with lower vaccine effectiveness estimation interval below 0%.
- Those with vaccine effectiveness point-estimate above 100% or below 0%.

**These analysis and plots are not suitable, since they are trying to portray vaccine effectiveness parametres that have been calculated on variables related to risk, not protection (as opposite to vaccine doses).**

```{r, eval = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
var_agegr_LASTIMMUN %>%
  filter(outcome == "hospitalisation") %>%
  filter(
    (original_model == "Before_60")
    | (original_model == "α_60")
    | (original_model == "Δ_Menor")
    | (original_model == "Δ_60")
    | original_model %in% c("Ό_60", "Ό_Menor")
  ) %>%
  filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(veffec_upper > 0 & veffec_upper <= 100) %>%
  filter(veffec > 0 & veffec < 100) %>%
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = veffec_lower, ymax = veffec_upper, colour = variant),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = veffec, colour = variant), size = 3) +
  facet_wrap(~agegroup, nrow = 2, strip.position = "left") +
  scale_colour_manual(values = colours) +
  # scale_y_continuous(limits = c(0, 100)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Vaccine effectiveness",
    x = "Time past since last exposure",
    colour = "Variant period",
    title = "Outcome: hospitalisation"
  )



var_agegr_LASTIMMUN %>%
  filter(outcome == "intensivecare") %>%
  filter(
    (original_model == "Before_60")
    | (original_model == "α_60")
    | (original_model == "Δ_Menor")
    | (original_model == "Δ_60")
    | original_model %in% c("Ό_60", "Ό_Menor")
  ) %>%
  filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(veffec_upper > 0 & veffec_upper <= 100) %>%
  filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = veffec_lower, ymax = veffec_upper, colour = variant),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = veffec, colour = variant), size = 3) +
  facet_wrap(~agegroup, nrow = 2, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 100)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Vaccine effectiveness",
    x = "Time past since last exposure",
    colour = "Variant period",
    title = "Outcome: ICU admission"
  )



var_agegr_LASTIMMUN %>%
  filter(outcome == "mortality") %>%
  filter(
    (original_model == "Before_60")
    | (original_model == "α_60")
    | (original_model == "Δ_Menor")
    | (original_model == "Δ_60")
    | original_model %in% c("Ό_60", "Ό_Menor")
  ) %>%
  filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(veffec_upper > 0 & veffec_upper <= 100) %>%
  filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = veffec_lower, ymax = veffec_upper, colour = variant),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = veffec, colour = variant), size = 3) +
  facet_wrap(~agegroup, nrow = 2, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 100)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Vaccine effectiveness",
    x = "Time past since last exposure",
    colour = "Variant period",
    title = "Outcome: 30-days mortality"
  )
```

### SUB-B2: NO MANUAL FILTERS Vaccine effectiveness by time since last exposure: infection/vaccination (segregated by variant period and age group)

Groups automatically excluded:

- Those with upper vaccine effectiveness estimation interval above 100%.
- Those with lower vaccine effectiveness estimation interval below 0%.
- Those with vaccine effectiveness point-estimate above 100% or below 0%.

**These analysis and plots are not suitable, since they are trying to portray vaccine effectiveness parametres that have been calculated on variables related to risk, not protection (as opposite to vaccine doses).**

```{r, eval = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
var_agegr_LASTIMMUN %>%
  filter(outcome == "hospitalisation") %>%
  # filter(
  #   (original_model == "Before_60")
  #   | (original_model == "α_60")
  #   | (original_model == "Δ_Menor")
  #   | (original_model == "Δ_60")
  #   | original_model %in% c("Ό_60", "Ό_Menor")
  # ) %>%
  filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(veffec_upper > 0 & veffec_upper <= 100) %>%
  filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = veffec_lower, ymax = veffec_upper, colour = variant),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = veffec, colour = variant), size = 3) +
  facet_wrap(~agegroup, nrow = 2, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 100)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Vaccine effectiveness",
    x = "Time past since last exposure",
    colour = "Variant period",
    title = "Outcome: hospitalisation"
  )



var_agegr_LASTIMMUN %>%
  filter(outcome == "intensivecare") %>%
  # filter(
  #   (original_model == "Before_60")
  #   | (original_model == "α_60")
  #   | (original_model == "Δ_Menor")
  #   | (original_model == "Δ_60")
  #   | original_model %in% c("Ό_60", "Ό_Menor")
  # ) %>%
  filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(veffec_upper > 0 & veffec_upper <= 100) %>%
  filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = veffec_lower, ymax = veffec_upper, colour = variant),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = veffec, colour = variant), size = 3) +
  facet_wrap(~agegroup, nrow = 2, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 100)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Vaccine effectiveness",
    x = "Time past since last exposure",
    colour = "Variant period",
    title = "Outcome: ICU admission"
  )



var_agegr_LASTIMMUN %>%
  filter(outcome == "mortality") %>%
  # filter(
  #   (original_model == "Before_60")
  #   | (original_model == "α_60")
  #   | (original_model == "Δ_Menor")
  #   | (original_model == "Δ_60")
  #   | original_model %in% c("Ό_60", "Ό_Menor")
  # ) %>%
  # filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(veffec_upper > 0 & veffec_upper <= 100) %>%
  filter(veffec > 0 & veffec < 100) %>%
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = veffec_lower, ymax = veffec_upper, colour = variant),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = veffec, colour = variant), size = 3) +
  facet_wrap(~agegroup, nrow = 2, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 100)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Vaccine effectiveness",
    x = "Time past since last exposure",
    colour = "Variant period",
    title = "Outcome: 30-days mortality"
  )
```

### SUB-B1 v2!: *Risk of infection* by time since last exposure: infection/vaccination (segregated by variant period and age group)

Groups manually excluded:

- All cases before alpha rise, younger than 60 years.
- All cases during alpha predominance, younger than 60 years.

Groups automatically excluded:

- Those with OR point-estimate above 10.

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
var_agegr_LASTIMMUN %>%
  filter(outcome == "hospitalisation") %>%
  filter(
    (original_model == "Before_60")
    | (original_model == "α_60")
    | (original_model == "Δ_Menor")
    | (original_model == "Δ_60")
    | original_model %in% c("Ό_60", "Ό_Menor")
  ) %>%
  # filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(OR_upper < 10) %>%
  # filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = OR_lower, ymax = OR_upper, colour = agegroup),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = OR, colour = agegroup), size = 3) +
  facet_wrap(~variant, nrow = 4, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 10.1)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Risk of infection (Odds Ratio)",
    x = "Time past since last exposure",
    colour = "Age group",
    title = "Outcome: hospitalisation"
  )



var_agegr_LASTIMMUN %>%
  filter(outcome == "intensivecare") %>%
  filter(
    (original_model == "Before_60")
    | (original_model == "α_60")
    | (original_model == "Δ_Menor")
    | (original_model == "Δ_60")
    | original_model %in% c("Ό_60", "Ό_Menor")
  ) %>%
  # filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(OR_upper < 10) %>%
  # filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = OR_lower, ymax = OR_upper, colour = agegroup),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = OR, colour = agegroup), size = 3) +
  facet_wrap(~variant, nrow = 4, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 10.1)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Risk of infection (Odds Ratio)",
    x = "Time past since last exposure",
    colour = "Age group",
    title = "Outcome: ICU admission"
  )



var_agegr_LASTIMMUN %>%
  filter(outcome == "mortality") %>%
  filter(
    (original_model == "Before_60")
    | (original_model == "α_60")
    | (original_model == "Δ_Menor")
    | (original_model == "Δ_60")
    | original_model %in% c("Ό_60", "Ό_Menor")
  ) %>%
  # filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(OR_upper < 10) %>%
  # filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = OR_lower, ymax = OR_upper, colour = agegroup),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = OR, colour = agegroup), size = 3) +
  facet_wrap(~variant, nrow = 4, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 10.1)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Risk of infection (Odds Ratio)",
    x = "Time past since last exposure",
    colour = "Age group",
    title = "Outcome: 30-days mortality"
  )
```

### SUB-B2 v2!: NO MANUAL FILTERS *Risk of infection* by time since last exposure: infection/vaccination (segregated by variant period and age groups)

Groups automatically excluded:

- Those with OR point-estimate above 10.

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
var_agegr_LASTIMMUN %>%
  filter(outcome == "hospitalisation") %>%
  # filter(
  #   (original_model == "Before_60")
  #   | (original_model == "α_60")
  #   | (original_model == "Δ_Menor")
  #   | (original_model == "Δ_60")
  #   | original_model %in% c("Ό_60", "Ό_Menor")
  # ) %>%
  # filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(OR_upper < 10) %>%
  # filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = OR_lower, ymax = OR_upper, colour = agegroup),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = OR, colour = agegroup), size = 3) +
  facet_wrap(~variant, nrow = 4, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 10.1)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Risk of infection (Odds Ratio)",
    x = "Time past since last exposure",
    colour = "Age group",
    title = "Outcome: hospitalisation"
  )



var_agegr_LASTIMMUN %>%
  filter(outcome == "intensivecare") %>%
  # filter(
  #   (original_model == "Before_60")
  #   | (original_model == "α_60")
  #   | (original_model == "Δ_Menor")
  #   | (original_model == "Δ_60")
  #   | original_model %in% c("Ό_60", "Ό_Menor")
  # ) %>%
  # filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(OR_upper < 10) %>%
  # filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = OR_lower, ymax = OR_upper, colour = agegroup),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = OR, colour = agegroup), size = 3) +
  facet_wrap(~variant, nrow = 4, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 10.1)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Risk of infection (Odds Ratio)",
    x = "Time past since last exposure",
    colour = "Age group",
    title = "Outcome: ICU admission"
  )



var_agegr_LASTIMMUN %>%
  filter(outcome == "mortality") %>%
  # filter(
  #   (original_model == "Before_60")
  #   | (original_model == "α_60")
  #   | (original_model == "Δ_Menor")
  #   | (original_model == "Δ_60")
  #   | original_model %in% c("Ό_60", "Ό_Menor")
  # ) %>%
  # filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(OR_upper < 10) %>%
  # filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = OR_lower, ymax = OR_upper, colour = agegroup),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = OR, colour = agegroup), size = 3) +
  facet_wrap(~variant, nrow = 4, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 10.1)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Risk of infection (Odds Ratio)",
    x = "Time past since last exposure",
    colour = "Age group",
    title = "Outcome: 30-days mortality"
  )
```

### Risk of infection by time since last exposure (segregated by number of vaccine doses and age groups)

Groups automatically excluded:

- Those with OR point-estimate above 10.

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
vacc_agegroup_LASTIMMUN %>%
  filter(outcome == "hospitalisation") %>% 
  # filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(OR_upper < 10) %>%
  # filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = OR_lower, ymax = OR_upper, colour = agegroup),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = OR, colour = agegroup), size = 3) +
  facet_wrap(~nvaccine, nrow = 4, strip.position = "left") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 10.1)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Risk of infection (Odds Ratio)",
    x = "Time past since last exposure (days)",
    colour = "Age group",
    title = "Outcome: hospitalisation"
  )



vacc_agegroup_LASTIMMUN %>%
  filter(outcome == "intensivecare") %>%
  # filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(OR_upper < 10) %>%
  # filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = OR_lower, ymax = OR_upper, colour = agegroup),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = OR, colour = agegroup), size = 3) +
  facet_wrap(~nvaccine, ncol = 4, strip.position = "top") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 10.1)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Risk of infection (Odds Ratio)",
    x = "Time past since last exposure (days)",
    colour = "Age group",
    title = "Outcome: ICU admission"
  )



vacc_agegroup_LASTIMMUN %>%
  filter(outcome == "mortality") %>%
  # filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(OR_upper < 10) %>%
  # filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = OR_lower, ymax = OR_upper, colour = agegroup),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = OR, colour = agegroup), size = 3) +
  facet_wrap(~nvaccine, ncol = 4, strip.position = "top") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 10.1)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Risk of infection (Odds Ratio)",
    x = "Time past since last exposure (days)",
    colour = "Age group",
    title = "Outcome: 30-days mortality"
  )
```

### ALTERNATIVE!! PREDICTOR: time past since last exposure (subset by nº of doses)

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
vacc_agegroup_LASTIMMUN %>%
  mutate(
    outcome = case_when(
      outcome == "hospitalisation" ~ "Hospital admission",
      outcome == "intensivecare" ~ "ICU admission",
      outcome == "intensivecare_only_hosp" ~ "ICU admission, only inpatient cases",
      outcome == "mortality" ~ "30-days mortality"
    ),
    outcome = factor(
      outcome,
      levels = rev(c(
        "Hospital admission", "ICU admission", "ICU admission, only inpatient cases", "30-days mortality"
      ))
    )
  ) %>%
  filter(outcome != "ICU admission, only inpatient cases") %>%
  # filter(veffec_lower > 0 & veffec_lower <= 100) %>%
  filter(OR_upper < 10) %>%
  # filter(veffec > 0 & veffec < 100) %>% 
  ggplot() +
  geom_errorbar(
    aes(x = predictor, ymin = OR_lower, ymax = OR_upper, colour = agegroup),
    linewidth = 1,
    width = 0.2
  ) +
  geom_point(aes(x = predictor, y = OR, colour = agegroup), size = 3) +
  facet_grid(rows = vars(outcome), cols = vars(nvaccine), switch = "y") +
  scale_colour_manual(values = colours) +
  scale_y_continuous(limits = c(0, 10.1)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  labs(
    y = "Risk of infection (Odds Ratio)",
    x = "Time past since last exposure (days)",
    colour = "Age group"
    # title = "Outcome: hospitalisation"
  )
```





# Upset Plot for describing categories

```{r cars}
# upset(data=vacunas, intersections = NOM_VACUNA,group.by = COD_NUHSA_ENCRIPTADO, order.by = "freq")
```

# Sankey Plots

## Data Load

You can also embed plots, for example:

```{r}


# Read final dataset file 
# estudio_vacunas_t_final_selected <-
#   readRDS(file = "/opt/datos_compartidos/vacunas_data/Rfiles/estudio_vacunas_t_final_selected.rds")
# nrow(estudio_vacunas_t_final_selected)
# summary(estudio_vacunas_t_final_selected$pautas_covid)


# Vaccinated and unvaccinated population before positive CODIV-19 infection by RT-PCR
# vacunados <-
#   estudio_vacunas_t_final_selected %>%
#   filter(pautas_covid %in% c("1", "2", "3", "4")) %>%
#   nrow()
# vacunados

# no_vacunados <-
#   estudio_vacunas_t_final_selected %>%
#   filter(pautas_covid %in% c("0")) %>%
#   nrow()
# no_vacunados

# Hospital admission data
# vacunados_ingresados <-
#   estudio_vacunas_t_final_selected %>%
#   filter(pautas_covid %in% c("1", "2", "3", "4")
#          & hosp_stay_bin == 1) %>%
#   nrow()
# vacunados_ingresados

# vacunados_no_ingresados <-
#   estudio_vacunas_t_final_selected %>%
#   filter(pautas_covid %in% c("1", "2", "3", "4")
#          & hosp_stay_bin == 0) %>%
#   nrow()
# vacunados_no_ingresados

# no_vacunados_ingresados <-
#   estudio_vacunas_t_final_selected %>%
#   filter(pautas_covid %in% c("0")
#          & hosp_stay_bin == 1) %>%
#   nrow()
# no_vacunados_ingresados

# no_vacunados_no_ingresados <-
#   estudio_vacunas_t_final_selected %>%
#   filter(pautas_covid %in% c("0")
#          & hosp_stay_bin == 0) %>%
#   nrow()
# no_vacunados_no_ingresados


# UCI admission data
# vacunados_ingresados_uci <-
#   estudio_vacunas_t_final_selected %>%
#   filter(pautas_covid %in% c("1", "2", "3", "4")
#          & icu_stay_bin == 1) %>%
#   nrow()
# vacunados_ingresados_uci

# vacunados_no_ingresados_uci <-
  # estudio_vacunas_t_final_selected %>%
  # filter(pautas_covid %in% c("1", "2", "3", "4")
         # & icu_stay_bin == 0) %>%
  # nrow()
# vacunados_no_ingresados_uci

# no_vacunados_ingresados_uci <-
  # estudio_vacunas_t_final_selected %>%
  # filter(pautas_covid %in% c("0")
         # & icu_stay_bin == 1) %>%
  # nrow()
# no_vacunados_ingresados_uci

# no_vacunados_no_ingresados_uci <-
  # estudio_vacunas_t_final_selected %>%
  # filter(pautas_covid %in% c("0")
         # & icu_sta/opt/datos_compartidos/vacunas_data/tablas/groups by nvac & lastimmun (logreg vaccines only, & variant)y_bin == 0) %>%
  # nrow()
# no_vacunados_no_ingresados_uci


# Mortality data
# vacunados_muertos <-
  # estudio_vacunas_t_final_selected %>%
  # filter(pautas_covid %in% c("1", "2", "3", "4")
         # & mortality30d == 1) %>%
  # nrow()
# vacunados_muertos

# vacunados_vivos <-
  # estudio_vacunas_t_final_selected %>%
  # filter(pautas_covid %in% c("1", "2", "3", "4")
         # & mortality30d == 0) %>%
  # nrow()
# vacunados_vivos

# no_vacunados_muertos <-
  # estudio_vacunas_t_final_selected %>%
  # filter(pautas_covid %in% c("0")
         # & mortality30d == 1) %>%
  # nrow()
# no_vacunados_muertos

# no_vacunados_vivos <-
  # estudio_vacunas_t_final_selected %>%
  # filter(pautas_covid %in% c("0")
         # & mortality30d == 0) %>%
  # nrow()
# no_vacunados_vivos

```
## Sankey Plot for Hospital Admission


```{r}
#Name of nodes
# A <- "Población de estudio"
# B <- "Vacunados"
# C <- "No Vacunados"
# D <- "No Hospitalizados"
# E <- "Hospitalizados"

# Make a connection data frame
# source <- c(A, A, B, B, C, C)
# target <- c(B, C, D, E, D, E)
# value <- c(vacunados, no_vacunados, vacunados_no_ingresados, vacunados_ingresados, no_vacunados_no_ingresados, no_vacunados_ingresados)
# data <- data.frame(source, target, value)

# From these flows we need to create a node data frame: it lists every entities involved in the flow
# nodes <-
#   data.frame("name" = c(
#     as.character(data$source),
#     as.character(data$target)
#     %>% unique()
#   ))

# With networkD3, connection must be provided using id, not using real name like in the links dataframe. So we need to reformat it.
# data$IDsource <- match(data$source, nodes$name) - 1
# data$IDtarget <- match(data$target, nodes$name) - 1


## Add link types for coloring purpose
# data$group = as.factor(c("type_0",
#                          "type_0",
#                          "type_1",
#                          "type_2",
#                          "type_1",
#                          "type_2"))

# Add a 'group' column to each node.
# nodes$group <-
#   as.factor(
#     c(
#       "node1",
#       "node1",
#       "node2",
#       "node2",
#       "node3",
#       "node3",
#       "node2",
#       "node3",
#       "node4",
#       "node5"
#     )
#   )

# Give a color for each group
# my_color <-
#   'd3.scaleOrdinal() .domain(["type_0", "type_1", "type_2", "node1", "node2", "node3", "node4", "node5"]) .range(["#a4b0be", "#7bed9f", "#ff6b81", "#747d8c", "#70a1ff", "#eccc68", "#2ed573", "#ff4757"])'

# Create the Sankey diagram
# SankeyDiagram <- sankeyNetwork(
#   Links = data,
#   Nodes = nodes,
#   Source = "IDsource",
#   Target = "IDtarget",
#   Value = "value",
#   NodeID = "name",
#   colourScale = my_color,
#   LinkGroup = "group",
#   NodeGroup = "group",
#   fontSize = 16,
#   fontFamily = 34,
#   nodeWidth = 100,
#   nodePadding = 20,
#   height = 400,
#   width = 1100,
#   iterations = 32,
#   sinksRight = FALSE
# )

# Print the Sankey diagram
# SankeyDiagram


```

## Sankey Plot for UCI Admission


```{r}
#Name of nodes
# A <- "Población de estudio"
# B <- "Vacunados"
# C <- "No Vacunados"
# G <- "No Ingresados en UCI"
# H <- "Ingresados en UCI"

# Make a connection data frame
# source <- c(A, A, B, B, C, C)
# target <- c(B, C, G, H, G, H)
# value <- c(vacunados, no_vacunados, vacunados_no_ingresados_uci, vacunados_ingresados_uci, no_vacunados_no_ingresados_uci, no_vacunados_ingresados_uci)
# data <- data.frame(source, target, value)

# From these flows we need to create a node data frame: it lists every entities involved in the flow
# nodes <-
#   data.frame("name" = c(
#     as.character(data$source),
#     as.character(data$target)
#     %>% unique()
#   ))

# With networkD3, connection must be provided using id, not using real name like in the links dataframe. So we need to reformat it.
# data$IDsource <- match(data$source, nodes$name) - 1
# data$IDtarget <- match(data$target, nodes$name) - 1


## Add link types for coloring purpose
# data$group = as.factor(c("type_0",
#                          "type_0",
#                          "type_1",
#                          "type_2",
#                          "type_1",
#                          "type_2"))

# Add a 'group' column to each node.
# nodes$group <-
#   as.factor(
#     c(
#       "node1",
#       "node1",
#       "node2",
#       "node2",
#       "node3",
#       "node3",
#       "node2",
#       "node3",
#       "node4",
#       "node5"
#     )
#   )

# Give a color for each group
# my_color <-
#   'd3.scaleOrdinal() .domain(["type_0", "type_1", "type_2", "node1", "node2", "node3", "node4", "node5"]) .range(["#a4b0be", "#7bed9f", "#ff6b81", "#747d8c", "#70a1ff", "#eccc68", "#2ed573", "#ff4757"])'

# Create the Sankey diagram
# SankeyDiagram <- sankeyNetwork(
#   Links = data,
#   Nodes = nodes,
#   Source = "IDsource",
#   Target = "IDtarget",
#   Value = "value",
#   NodeID = "name",
#   colourScale = my_color,
#   LinkGroup = "group",
#   NodeGroup = "group",
#   fontSize = 16,
#   fontFamily = 34,
#   nodeWidth = 100,
#   nodePadding = 20,
#   height = 400,
#   width = 1100,
#   iterations = 32,
#   sinksRight = FALSE
# )

# Print the Sankey diagram
# SankeyDiagram

```

## Sankey Plot for Survival/Mortality


```{r}
#Name of nodes
# A <- "Población de estudio"
# B <- "Vacunados"
# C <- "No Vacunados"
# I <- "Supervivientes"
# J <- "Fallecidos"

# Make a connection data frame
# source <- c(A, A, B, B, C, C)
# target <- c(B, C, I, J, I, J)
# value <- c(vacunados, no_vacunados, vacunados_vivos, vacunados_muertos, no_vacunados_vivos, no_vacunados_muertos)
# data <- data.frame(source, target, value)

# From these flows we need to create a node data frame: it lists every entities involved in the flow
# nodes <-
#   data.frame("name" = c(
#     as.character(data$source),
#     as.character(data$target)
#     %>% unique()
#   ))

# With networkD3, connection must be provided using id, not using real name like in the links dataframe. So we need to reformat it.
# data$IDsource <- match(data$source, nodes$name) - 1
# data$IDtarget <- match(data$target, nodes$name) - 1


## Add link types for coloring purpose
# data$group = as.factor(c("type_0",
#                          "type_0",
#                          "type_1",
#                          "type_2",
#                          "type_1",
#                          "type_2"))

# Add a 'group' column to each node.
# nodes$group <-
#   as.factor(
#     c(
#       "node1",
#       "node1",
#       "node2",
#       "node2",
#       "node3",
#       "node3",
#       "node2",
#       "node3",
#       "node4",
#       "node5"
#     )
#   )

# Give a color for each group
# my_color <-
#   'd3.scaleOrdinal() .domain(["type_0", "type_1", "type_2", "node1", "node2", "node3", "node4", "node5"]) .range(["#a4b0be", "#7bed9f", "#ff6b81", "#747d8c", "#70a1ff", "#eccc68", "#2ed573", "#ff4757"])'

# Create the Sankey diagram
# SankeyDiagram <- sankeyNetwork(
#   Links = data,
#   Nodes = nodes,
#   Source = "IDsource",
#   Target = "IDtarget",
#   Value = "value",
#   NodeID = "name",
#   colourScale = my_color,
#   LinkGroup = "group",
#   NodeGroup = "group",
#   fontSize = 16,
#   fontFamily = 34,
#   nodeWidth = 100,
#   nodePadding = 20,
#   height = 400,
#   width = 1100,
#   iterations = 32,
#   sinksRight = FALSE
# )

# Print the Sankey diagram
# SankeyDiagram
```


# Rentabilidad Edad

```{r}
# Definir los parámetros
# costo_hospitalizacion <- 4612.76
# precio_vacuna <- 31
# dosis_equiv <- 148
# nnt <- 148
# red_abs_riesgo <- 0.0068

# Definir un vector de edades de 18 a 120
# edades <- 18:120

# Calcular la rentabilidad para cada edad
# rentabilidad <- (costo_hospitalizacion/dosis_equiv) / nnt / (1 - red_abs_riesgo)

# Crear un dataframe con los resultados
# df <- data.frame(edad = edades, rentabilidad = rentabilidad)

```

